---
title: "English-hearing toddlers learn words from consistent and variable labels"
shorttitle: "Learning from consistent and variable labels"
author:
  - name: Kennedy Casey
    corresponding: true
    orcid: 0000-0001-5326-3080
    email: kcasey@princeton.edu
    affiliations:
      - name: Princeton University
        department: Department of Psychology
        address: 40 Woodlands Way
        city: Princeton
        region: NJ
        country: USA
        postal-code: 08544
  - name: Casey Lew-Williams
    orcid: 0000-0002-8781-4458
    email: caseylw@princeton.edu
    affiliations:
      - name: Princeton University
        department: Department of Psychology
        address: 40 Woodlands Way
        city: Princeton
        region: NJ
        country: USA
        postal-code: 08544
author-note:
  disclosures:
    study-registration: "The study design and analysis plan was pre-registered ([https://osf.io/uqz6d](https://osf.io/uqz6d))"
    data-sharing: "Anonymized data are available at LINK"
    conflict-of-interest: "The authors declare no conflicts of interest"
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "This document is a template."
keywords: [word learning, child-directed speech, wordplay, eye-tracking]
bibliography: LexVar.bib
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
execute: 
  echo: false
  warning: false
  message: false
  error: false
  eval: true

---
```{r setup}
library(tidyverse)
library(readxl)
library(english)
library(tools)
library(flextable)
library(brms)
library(loo)

# reprocess raw data? set to TRUE or FALSE
reprocess = FALSE
if (reprocess) {
  source("analysis/process.R")
}

# set condition colors
condition.colors <- c("stable" = "#ef5039", 
                      "unnaturalv" = "#04a48c", 
                      "naturalv" = "#3f5eaa")
```

# Method
This study was pre-registered ([https://osf.io/uqz6d](https://osf.io/uqz6d)). All stimuli, anonymized data, and analysis scripts are publicly available at LINK.

```{r, echo=FALSE}
log <- read_xlsx("data/metadata/ANON-ptcp-log.xlsx")
final.sample <- filter(log, usable_checked == "y") 
usable <- final.sample %>% nrow()
unusable <- filter(log, usable_checked == "n") %>% nrow()
unusable.cal <- filter(log, str_detect(exclusion_reason, "calibration")) %>% nrow()
unusable.data <- filter(log, str_detect(exclusion_reason, "trial")) %>% nrow()
unusable.inter <- filter(log, str_detect(exclusion_reason, "interruption")) %>% nrow()

ethnicity <- final.sample %>%
  mutate(ethnicity = case_when(
    ethnicity %in% c("black", "asian", "not reported") ~ str_to_sentence(ethnicity), 
    ethnicity == "white" ~ "Non-Latinx white", 
    ethnicity == "white/hispanic" ~ "Latinx white", 
    ethnicity %in% c("white/black", "asian/white", "mixed") ~ "Mixed race"
  )) %>%
  group_by(ethnicity) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = round(n/sum(n)*100, 2)) %>%
  arrange(desc(prop))

ages <- final.sample %>%
  transmute(id = str_extract(ptcp_id, "[0-9][0-9][0-9]"), 
            age_c = datawizard::standardize(age))
```

## Participants
Participants (_N_ = `r usable`), aged 18 to 36 months (*M*~age~ = `r format(round(mean(final.sample$age), 2), nsmall = 2)`), were full-term, monolingual English-hearing children (>80% reported exposure) with no known hearing, vision, language, and/or developmental delays. Caregivers self-reported race/ethnicity information for their children in an open-ended format when they initially signed up for our volunteer database. `r sub("^([a-z])", "\\U\\1", as.character(english(ethnicity[2,]$n)), perl = TRUE)` caregivers  (`r ethnicity[2,]$prop`% of the final sample) chose not to report this information. Participants were primarily `r ethnicity[1,]$ethnicity` (_N_ = `r ethnicity[1,]$n`, `r ethnicity[1,]$prop`%) with some representation of other races/ethnicities: Mixed race (_N_ = `r ethnicity[2,]$n`, `r ethnicity[2,]$prop`%), Asian (_N_ = `r ethnicity[4,]$n`, `r ethnicity[4,]$prop`%), Black (_N_ = `r ethnicity[5,]$n`, `r ethnicity[5,]$prop`%), `r ethnicity[6,]$n` (_N_ = `r ethnicity[6,]$n`, `r ethnicity[6,]$prop`%).

The final sample size was pre-registered and determined using an a priori power analysis conducted in G*Power [@faul2009statistical] and informed by prior eye-tracking studies with a similar age range [@nencheva2021moment; @quam2023protracted]. `r sub("^([a-z])", "\\U\\1", as.character(english(unusable)), perl = TRUE)` additional children participated but were excluded from the final analyses due to refusal to wear sticker or complete calibration (_N_ = `r unusable.cal`), significant interruption by family member (_N_ = `r unusable.inter`), or contribution of too few usable trials based on pre-registered criteria (_N_ = `r unusable.data`).

## Stimuli
Auditory stimuli were recorded by female native English speaker using naturalistic child-directed prosody in a sound-attenuated room. Each target word (see Table 1) appeared in four different sentence frames for exposure trials ("Look, _____! That's a _____. Do you see the ______? Yes, that's a ______.") and in two different sentence frames for test trials ("Can you find the _____? See the _____?"). The exact same sentence frame recordings were used across all target words. Trials were created in Praat [CITE] by concatenating each target word onto the final position of each sentence frame. To avoid co-articulation artifacts, in trials with articles preceding the target word, the target words were spliced along with the article. [ADD TIMING INFO]. All auditory stimuli were normed to a standard mean intensity of 65dB. Minor acoustic edits were made on some sentences to ensure that all trials sounded naturally produced and had comparable loudness and pitch.

```{r stimuli-table}
tribble(
  ~Standard, ~`Reduplicated (Typical)`, ~`Reduplicated (Atypical)`, ~`Diminutivized (Typical)`, ~`Diminutivized (Atypical)`,
  "tanzer\n/tæn.zɚ/", "tan-tan\n/tæn.tæn/", "zer-zer\n/zɚ.zɚ/", "tanny\n/tæn.i/", "y-zer\n/i.zɚ/",
  "bogdum\n/bɑg.dʌm/", "bog-bog\n/bɑg.bɑg/", "dum-dum\n/dʌm.dʌm/", "boggy\n/bɑg.i/", "y-dum\n/i.dʌm/",
  "kaisol\n/kaɪ.sɔl/", "kai-kai\n/kaɪ.kaɪ/", "sol-sol\n/sɔl.sɔl/", "kaisy\n/kaɪ.si/", "y-sol\n/i.sɔl/",
  "mobet\n/moʊ.bɛt/", "mo-mo\n/moʊ.moʊ/", "bet-bet\n/bɛt.bɛt/", "moby\n/moʊ.bi/", "y-bet\n/i.bɛt/",
  "woifer\n/wɔɪ.fɚ/", "woi-woi\n/wɔɪ.wɔɪ/", "fer-fer\n/fɚ.fɚ/", "woify\n/wɔɪ.fi/", "y-fer\n/i.fɚ/",
  "voopin\n/vu.pɪn/", "voo-voo\n/vu.vu/", "pin-pin\n/pɪn.pɪn/", "voopy\n/vu.pi/", "y-pin\n/i.pɪn/"
) %>% flextable() %>%
  set_caption("Table 1. Target word forms.") %>%
  set_table_properties(width = 1, layout = "autofit") %>%
  fontsize(size = 10, part = "all") %>%
  align(align = "left", part = "all")
```

The visual stimuli consisted of six isoluminant objects with 12 unique colors (two per object) that were modeled after @nencheva2021moment's stimuli and hand-drawn using Procreate software [@procreate].

## Procedure
Toddlers sat on their caregiver's lap, approximately 60cm away from a 17-inch monitor equipped with an EyeLink 1000 Plus eye-tracker, in a dimly-lit testing room. Caregivers wore opague sunglasses to prevent them from influencing their children's behavior. We used a 3-point calibration procedure along the horizontal dimension (x-axis) of the screen. The experiment consisted of three blocks (1 per condition). Each block consisted of two phases: an exposure phase, during which participants learned novel word-object pairings, and a test phase, during which toddlers were immediately tested on their recognition of the word-object mappings. An isoluminant, non-linguistic attention-getting video was shown after every block.

During the exposure phase (4 trials per block), children saw a single target object appear at the center of the screen. Before the onset of each trial, children saw a blank gray screen. Each trial started after a short, variable-length ITI (min: 250ms, max: 750ms, mean: 500ms) or when the participant looked at the screen. Next, the target object appeared silently on the screen for 1500ms or until the participant looked at the screen. Then, the exposure audio played for XXXX-XXXXms (_M_ = XXXXms). After the offset of the audio, the target object remained on the screen for an additional 1500ms.

During the test phase (4 trials per block), toddlers saw the two objects from the preceding exposure phase side-by-side on the screen. Each trial started with an ITI (same structure as training phase), followed by a 2000ms silent period with both objects on the screen.

## Conditions
Participants learned words in three conditions (within-subjects manipulation). In all conditions, participants learned the names of
two different novel objects across four alternating exposure trials (2 per word). In the _consistent_ condition, participants heard the standard form of the target word repeated 5x per trial (e.g., "Look, tanzer! That's a tanzer...tanzer. Do you see the tanzer? Yes, that's a tanzer."). In the _typical variation_ condition, participants heard the standard form of the target word repeated 2x per trial in addition to one reduplicated form and one diminutivized form per trial (e.g., "Look, tanzer! That's a tan-tan. Do you see the tanny? Yes, that's a tanzer."). Critically, the number of exposures to the first syllable of the target word (i.e., "tan") are equivalent across these two conditions. In the _atypical variation_ (control) condition, participants heard the standard form of the target word repeated 2x per trial in addition to 1x exposure each of reversed versions of the typical reduplicated and diminutivized forms (e.g., "Look, tanzer! That's a zer-zer. Do you see the y-zer? Yes, that's a tanzer."). Participants were assigned to one of 18 pseudorandomized orders such that the order of presentation of blocks and the assignment of target labels to target objects was counterbalanced across conditions.

```{r usable-stats}
path <- "data/processed/usable/"
usable <- tibble()
for (i in list.files(path)) {
  usable <- bind_rows(usable, 
                      read_csv(paste0(path, i), show_col_types = FALSE))
}

n.conditions <- usable %>%
  group_by(id) %>%
  summarize(n = length(unique(condition))) %>%
  ungroup() %>%
  summarize(mean = round(mean(n), 2),
            median = round(median(n), 2),
            min = min(n),
            max = max(n))

n.trials <- usable %>%
  group_by(id) %>%
  summarize(n = length(unique(trial))) %>%
  ungroup() %>%
  summarize(mean = round(mean(n), 2),
            median = round(median(n), 2),
            min = min(n),
            max = max(n))
```
## Data pre-processing
Gaze behavior was used to determine toddlers' recognition of novel word-object pairings during two critical windows of analysis (300ms-2000ms following each target word onset; CITE). We calculated participants' proportion looking to target objects during test trials as the measure of word-recognition accuracy. For any given test trial to be included in analyses, children must have looked to the target and/or distractor object at least once prior to the onset of the label and for at least 1/3 of each critical post-labeling window. In order for a block to be included, children must have contributed usable data for at least 2 of 4 test trials. Finally, in order for a participant to be included, they must have contributed at least 2 of 3 blocks. On average, participants contributed `r n.trials$mean` trials (median = `r n.trials$median`, range = `r n.trials$min`--`r n.trials$max`) across `r n.conditions$mean` conditions (median = `r n.conditions$median`).

Following @nencheva2021moment, toddlers' raw pupil size values (automatically recorded by the EyeLink eye-tracker) were first converted to percent change from the average pupil size of the trial. ADD MORE DETAILS AFTER CONFIRMING ANALYSES BELOW. <!--directly from mira's paper, need to incorporate in code and edit wording accordingly: In order to identify artifacts due to partial blinks, variation in pupil size was estimated with a sliding 0.05-s window. Any 0.05-s intervals with a range of change exceeding 15% were removed as artifacts (Merritt, Keegan, & Mercer, 1994). To estimate the temporal alignment of pupil responses between participants, we transformed a distance measure (DTW, the same measure used for computing distance between pitch contours in
Study 1) quantifying the dissimilarity between two time series, into synchrony, an inversely related measure of the similarity between two time series. Because the analysis time window varied, we used a DTW distance measure that was normalized for duration, which eliminated effects of duration differences on synchrony. To convert from distance to synchrony, we subtracted the distance results from the maximum distance between two participants on a trial in the current experiment. To compare synchrony for CDS and ADS, we computed pairwise synchrony over the pupil size time window spanning the three 15-s story segments for all possible pairs of participants. We excluded segments in which more than half of the data were missing, and interpolated gaps smaller than 100 ms using Stineman interpolation (Johannesson, Bjornsson, Grothendieck, & Johannesson, 2018), which yielded the most accurate interpolation of all the techniques we tested (see Figure S4).--->

# Results
## Word recognition measured by gaze behavior
```{r test-prep}
rewrite = TRUE
path <- "data/processed/"

if (rewrite) {
  # test.props <- tibble()
  # test.props.bywindow <- tibble()
  # frames <- tibble()
  frames.cbp <- tibble()
  
  included.files <- list.files(path, "-test.csv")
  
  for (i in included.files) {
    
    curr.file <- read_csv(paste0(path, i))
    
    id <- str_extract(unique(curr.file$pid), "[0-9][0-9][0-9]")
    
    status <- filter(log, str_detect(ptcp_id, id))$usable_checked
    
    if (status != "n" & !is.na(status)) {
      # read in usable trials from pre-processing stage
      usable.trials <- read_csv(paste0("data/processed/usable/", id, ".csv")) %>%
        pull(trial)
      
      if (length(usable.trials) > 0) {
        
      curr.file.cbp <- curr.file %>% 
        select(pid, age_m, order, trial, x, pupil, base_word:window) %>%
        filter(trial %in% usable.trials) %>%
        group_by(pid, trial, condition, fix, base_word, window) %>%
        filter(length(unique(frame)) >= (2000-300)/3) %>%
        ungroup()
        
      # curr.file <- curr.file %>%
      #   select(pid, age_m, order, trial, x, pupil, base_word:window) %>%
      #   filter(trial %in% usable.trials & !is.na(window)) %>%
      #   group_by(pid, trial, condition, fix, base_word, window) %>%
      #   filter(length(unique(frame)) >= (2000-300)/3) %>%
      #   ungroup()
      # 
      # curr.test.props <- curr.file %>%
      #   group_by(pid, trial, condition, fix, base_word) %>%
      #   summarize(total = n()) %>%
      #   pivot_wider(names_from = "fix", values_from = "total", 
      #               id_cols = c(pid, trial, condition, base_word))
      # 
      # curr.test.props.bywindow <- curr.file %>%
      #   group_by(pid, trial, condition, fix, base_word, window) %>%
      #   summarize(total = n()) %>%
      #   pivot_wider(names_from = "fix", values_from = "total", 
      #               id_cols = c(pid, trial, condition, base_word, window))
      # 
      # if ("target" %in% colnames(curr.test.props)) {
      #   curr.test.props <- curr.test.props %>%
      #     mutate(target = ifelse(is.na(target), 0, target))
      #   
      # } else {
      #   curr.test.props <- curr.test.props %>%
      #     mutate(target = 0)
      # }
      # 
      # if ("distracter" %in% colnames(curr.test.props)) {
      #   curr.test.props <- curr.test.props %>%
      #     mutate(distracter = ifelse(is.na(distracter), 0, distracter))
      # } else {
      #   curr.test.props <- curr.test.props %>%
      #     mutate(distracter = 0)
      # }
      # 
      # curr.test.props <- curr.test.props %>%
      #   mutate(total = target + distracter, 
      #          prop = target/total)
      # 
      # test.props <- bind_rows(test.props, curr.test.props)
      # 
      # if ("target" %in% colnames(curr.test.props.bywindow)) {
      #   curr.test.props.bywindow <- curr.test.props.bywindow %>%
      #     mutate(target = ifelse(is.na(target), 0, target))
      #   
      # } else {
      #   curr.test.props.bywindow <- curr.test.props.bywindow %>%
      #     mutate(target = 0)
      # }
      # 
      # if ("distracter" %in% colnames(curr.test.props)) {
      #   curr.test.props.bywindow <- curr.test.props.bywindow %>%
      #     mutate(distracter = ifelse(is.na(distracter), 0, distracter))
      # } else {
      #   curr.test.props.bywindow <- curr.test.props.bywindow %>%
      #     mutate(distracter = 0)
      # }
      # 
      # curr.test.props.bywindow <- curr.test.props.bywindow %>%
      #   mutate(total = target + distracter, 
      #          prop = target/total)
      # 
      # test.props.bywindow <- bind_rows(test.props.bywindow, curr.test.props.bywindow)
      # 
      # curr.frames <- curr.file %>%
      #   select(pid, age_m, order, trial, x, pupil, base_word:window) %>%
      #   filter(trial %in% usable.trials) %>%
      #   mutate(fix_num = case_when(
      #     fix == "distracter" ~ 0, 
      #     fix == "target" ~ 1, 
      #     TRUE ~ NA
      #   ))
      
      curr.frames.cbp <- curr.file.cbp %>%
        select(pid, age_m, order, trial, x, pupil, base_word:window) %>%
        filter(trial %in% usable.trials) %>%
        mutate(fix_num = case_when(
          fix == "distracter" ~ 0, 
          fix == "target" ~ 1, 
          TRUE ~ NA
        ))
      
      # frames <- bind_rows(frames, curr.frames)
      frames.cbp <- bind_rows(frames.cbp, curr.frames.cbp)
      
      }
    }
  }
  
  # export proportion looking data
  
  # write_csv(test.props, "data/test-props.csv")
  
  # write_csv(test.props.bywindow, "data/test-props-bywindow.csv")
  
  # export full timecourse data
    # data <- frames %>%
    #   mutate(time = frame, 
    #          condition = factor(condition, levels = c("unnaturalv", "stable", "naturalv"))) %>%
    #   select(pid, age_m, order, trial, condition, base_word, pupil, time, fix_num, window,
    #          first_label, second_label) %>%
    #   group_by(pid) %>%
    #   mutate(block_order = list(unique(condition))) %>%
    #   ungroup() %>%
    #   mutate(condition_order = map2_int(block_order, condition, ~which(.x == .y)))
    # write_csv(data, "data/prepped-data.csv")
    
    # export eyetrackingR version of data
    eyetrackingr.data <- frames.cbp %>%
      mutate(time = frame, 
             condition = factor(condition, levels = c("unnaturalv", "stable", "naturalv"))) %>%
      mutate(target = ifelse(fix_num == 1, 1, 0),
             distracter = ifelse(fix_num == 0, 1, 0)) %>%
      select(-fix_num)
    write_csv(eyetrackingr.data, "data/eyetrackingr-data.csv")
} else {
  data <- read_csv("data/prepped-data.csv")
  test.props <- read_csv("data/test-props.csv")
  test.props.bywindow <- read_csv("data/test-props-bywindow.csv")
  eyetrackingr.data <- read_csv("data/eyetrackingr-data.csv")
}
```

```{r test-bayes, results = "hide"}
model.data <- test.props %>%
  mutate(id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id") %>%
  mutate(condition = relevel(as.factor(condition), ref = "unnaturalv"))

model.data.bywindow <- test.props.bywindow %>%
  mutate(id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id") %>%
  mutate(condition = relevel(as.factor(condition), ref = "unnaturalv"))

rerun = FALSE
if (rerun) {
  m.base <- brm(
  bf(prop ~ condition + (1 + condition | id) + (1 + condition | base_word),
     phi ~ 1,  # precision
     zoi ~ 1,  # zero-one inflation
     coi ~ 1), # conditional one-inflation (if needed)
  data = model.data,
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 4000,
  control = list(adapt_delta = 0.99)
  )
  saveRDS(m.base, file = "models/m-base.rds")
  
  m.age <- brm(
  bf(prop ~ condition + age_c + (1 + condition | id) + (1 + condition + age_c | base_word),
     phi ~ 1,  # precision
     zoi ~ 1,  # zero-one inflation
     coi ~ 1), # conditional one-inflation (if needed)
  data = model.data,
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 4000,
  control = list(adapt_delta = 0.99)
  )
  saveRDS(m.age, file = "models/m-age.rds")
  
  m.age.int <- brm(
  bf(prop ~ condition * age_c + (1 + condition | id) + (1 + condition * age_c | base_word),
     phi ~ 1,  # precision
     zoi ~ 1,  # zero-one inflation
     coi ~ 1), # conditional one-inflation (if needed)
  data = model.data,
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 4000,
  control = list(adapt_delta = 0.99)
  )
  saveRDS(m.age.int, file = "models/m-age-int.rds")
  
  m.window <- brm(
  bf(prop ~ condition + age_c + window + (1 + condition  + window | id) + (1 + condition + window + age_c | base_word),
     phi ~ 1,  # precision
     zoi ~ 1,  # zero-one inflation
     coi ~ 1), # conditional one-inflation (if needed)
  data = model.data.bywindow,
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 4000,
  control = list(adapt_delta = 0.99)
)
  saveRDS(m.window, file = "models/m-window.rds")
  
  m.window1 <- brm(
  bf(prop ~ condition + age_c + (1 + condition | id) + (1 + condition + age_c | base_word),
     phi ~ 1,  # precision
     zoi ~ 1,  # zero-one inflation
     coi ~ 1), # conditional one-inflation (if needed)
  data = filter(model.data.bywindow, window == "first"),
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 4000,
  control = list(adapt_delta = 0.99)
)
  saveRDS(m.window1, file = "models/m-window1.rds")
  
  m.window2 <- brm(
  bf(prop ~ condition + age_c + (1 + condition | id) + (1 + condition + age_c | base_word),
     phi ~ 1,  # precision
     zoi ~ 1,  # zero-one inflation
     coi ~ 1), # conditional one-inflation (if needed)
  data = filter(model.data.bywindow, window == "second"),
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 4000,
  control = list(adapt_delta = 0.99)
)
  saveRDS(m.window2, file = "models/m-window2.rds")
} else {
  m.base <- readRDS("models/m-base.rds")
  m.age <- readRDS("models/m-age.rds")
  m.age.int <- readRDS("models/m-age-int.rds")
  m.window <- readRDS("models/m-window.rds")
  m.window1 <- readRDS("models/m-window1.rds")
  m.window2 <- readRDS("models/m-window2.rds")
}


# no clear evidence that adding interaction term improves model fit
loo(m.age, m.age.int)


# check how well predictions match observed data
#pp_check(m.base, type = "intervals")

# summarize posterior
posterior_summary(m.age, prob = 0.89)
conditional_effects(m.age)

# generate posterior predictions per condition
newdata <- data.frame(
  condition = factor(c("unnaturalv", "stable", "naturalv"), levels = levels(model.data$condition)), 
  age_c = 0
)

# get fitted estimates of the mean (mu parameter of the beta)
preds <- posterior_epred(m.age, newdata = newdata, re_formula = NA)

# calculate means and 89% credible intervals
pred_summary <- as.data.frame(preds) %>%
  setNames(levels(newdata$condition)) %>%
  pivot_longer(everything(), names_to = "condition", values_to = "mu") %>%
  group_by(condition) %>%
  summarize(
    mean = mean(mu),
    ci_lower = quantile(mu, 0.055),
    ci_upper = quantile(mu, 0.945)
  )

print(pred_summary)
```



```{r extract-summary, echo=FALSE}
library(brms)
library(dplyr)
library(tibble)

# Extract posterior summaries
post_sum <- posterior_summary(m.age, prob = 0.89)

# Pull out fixed effects
natv_est <- post_sum["b_conditionnaturalv", "Estimate"]
natv_sd <- post_sum["b_conditionnaturalv", "Est.Error"]
natv_lb <- natv_est - qnorm(0.945) * natv_sd
natv_ub <- natv_est + qnorm(0.945) * natv_sd

stable_est <- post_sum["b_conditionstable", "Estimate"]
stable_sd <- post_sum["b_conditionstable", "Est.Error"]
stable_lb <- stable_est - qnorm(0.945) * stable_sd
stable_ub <- stable_est + qnorm(0.945) * stable_sd

age_est <- post_sum["b_age_c", "Estimate"]
age_sd <- post_sum["b_age_c", "Est.Error"]
age_lb <- age_est - qnorm(0.945) * age_sd
age_ub <- age_est + qnorm(0.945) * age_sd

# Posterior probabilities
draws <- as_draws_df(m.age)
prob_natv <- mean(draws$b_conditionnaturalv > 0)
prob_stable <- mean(draws$b_conditionstable > 0)
prob_age <- mean(draws$b_age_c > 0)
```

We fit a Bayesian zero-one inflated beta regression model [CITE] to examine the effects of input condition and age on children's accuracy of novel word recognition (i.e., proportion looking to the target object), using the `brms` package [CITE]. The model included fixed effects for condition (atypical variation vs. typical variation vs. consistent) and mean-centered age, along with the maximal random effects structure (by-participant and by-item random slopes and intercepts). Following our pre-registered analysis plan, we tested whether including an interaction between condition and age improved model fit---we compared the model with the interaction term to the simpler additive model using approximate leave-one-out cross-validation [CITE]. This analysis suggested that the difference in predictive accuracy was small, providing no strong evidence that the interaction term improved model fit (ΔELPD = `r round(loo_compare(loo(m.age), loo(m.age.int))[1, "elpd_diff"], 2)`, SE = `r round(loo_compare(loo(m.age), loo(m.age.int))[1, "se_diff"], 2)`).

Condition effects were weakly positive but uncertain. For the typical variation condition, the estimated effect was `r round(natv_est, 3)`, SD = `r round(natv_sd, 3)`, with a 89% credible interval from `r round(natv_lb, 3)` to `r round(natv_ub, 3)`. For the consistent condition, the estimate was `r round(stable_est, 3)`, SD = `r round(stable_sd, 3)`, 89% CI [`r round(stable_lb, 3)`, r `round(stable_ub, 3)`]. The effect of age was small (`r round(age_est, 3)`, SD = `r round(age_sd, 3)`), 89% CI [`r round(age_lb, 3)`, `r round(age_ub, 3)`].

Posterior probabilities that each effect was greater than zero were: `r round(prob_natv, 2)` for the effect of the typical variation condition (relative to atypical variation condition), `r round(prob_stable, 2)` for the effect of the consistent condition (relative to the atypical variation condition), and `r round(prob_age, 2)` for the effect of age. These results suggest limited evidence that either input condition or age significantly affected performance, though the posterior distributions for all three predictors leaned slightly positive.

```{r fixed_effects_summary, echo=FALSE}
library(bayesplot)
library(tidybayes)

# Extract fixed effects
fixef_summary <- posterior_summary(m.age, prob = 0.89)[c("b_conditionnaturalv", "b_conditionstable", "b_age_c"), ]
knitr::kable(round(fixef_summary, 3), caption = "Fixed effects (posterior mean ± SD with 89% credible intervals).")
```

```{r fig1}
#| label: fig1
#| fig-cap: "Figure 1: Timecourse of participants' looks during test trials with shaded standard error regions. The gray dashed line at 0.5 represents chance looking. The black dashed lines indicate the critical analysis window (300-2000ms) following target word onset."
#| width: 12
#| height: 6

data %>% 
  mutate(accuracy = fix_num) %>%
  filter(time >= -1500 & time <= 6000) %>%
  group_by(pid, time, condition, trial) %>%
  summarize(accuracy = mean(accuracy, na.rm = TRUE)) %>% 
  group_by(time, condition) %>% # aggregate across participants
  summarize(
    N = sum(!is.na(accuracy)),
    SD = sd(accuracy, na.rm = TRUE),
    SE = SD/sqrt(N),
    accuracy = mean(accuracy, na.rm = TRUE),
    lower = accuracy-SE,
    upper = accuracy+SE
  ) %>%
  ggplot(aes(x=time, y=accuracy, color = condition)) + 
  scale_fill_manual(values = condition.colors, labels = c("Atypical variation", "Consistent", "Typical variation")) +
  scale_color_manual(values = condition.colors, labels = c("Atypical variation", "Consistent", "Typical variation")) +
  geom_hline(aes(yintercept=0.5),linetype="dashed",color='gray') + 
  geom_smooth(aes(ymin=lower, ymax=upper, fill = condition), stat="identity") +
  geom_vline(xintercept = 300, linetype = "dashed", color = "black") +
  geom_vline(xintercept = 2000, linetype = "dashed", color = "black") +
  geom_line() +
  theme_test(base_size=15) +
  coord_cartesian(xlim=c(-500,2500), ylim= c(-.01,1.01),expand=F) +
  scale_x_continuous(breaks=seq(from=-500,to=2500,by=500)) +
  labs(x="Time from target word onset onset (ms)",
       y="Accuracy\n(Proportion looking to target object)", 
       color = "Condition", 
       fill = "Condition") +
  theme(axis.title = element_text(face = "bold"), 
        legend.title = element_text(face = "bold"))
```

```{r test2}
library(lme4)
library(lmerTest)
m <- lmer(prop ~ condition + age_c + (1 + condition | id), 
          data = model.data)

library(glmmTMB)
model.data <- model.data %>%
  mutate(prop_beta = (prop * (nrow(model.data) - 1) + 0.5) / nrow(model.data))

m_beta <- glmmTMB(
  prop_beta ~ condition + 
    (1 | id) + 
    (1 | base_word),
  data = model.data,
  family = beta_family()
)

bypid.bywindow.condition.summary <- data %>%
  filter(!is.na(window)) %>%
  mutate(condition = factor(condition, levels = c("stable", "naturalv", "unnaturalv")), 
         window = factor(window, levels = c("first", "second"), 
                         labels = c("post-naming window 1", 
                                    "post-naming window 2"))) %>%
  group_by(pid, condition, window, age_m) %>%
  summarize(accuracy = mean(fix_num)) %>%
  ungroup()

data %>%
  filter(!is.na(window)) %>%
  mutate(condition = factor(condition, levels = c("stable", "naturalv", "unnaturalv")), 
         window = factor(window, levels = c("first", "second"), 
                         labels = c("post-naming window 1", 
                                    "post-naming window 2"))) %>%
  group_by(pid, condition, window) %>%
  summarize(accuracy = mean(fix_num)) %>%
  ungroup() %>%
  group_by(condition, window) %>%
  summarize(mean = mean(accuracy), 
            sd = sd(accuracy), 
            ci = 1.96*sd/sqrt(n())) %>%
  ggplot(aes(x = condition, y = mean, color = condition, fill = condition)) + 
  facet_grid(.~window) +
  geom_hline(yintercept = 0.5, color = "gray", linetype = "dashed", size = 1.25) + 
  geom_bar(stat = "identity", alpha = 0.6, size = 2) + 
  geom_jitter(data = bypid.bywindow.condition.summary, 
              mapping = aes(x = condition, y = accuracy, color = condition), 
              width = 0.1, size = 3) + 
  geom_errorbar(aes(ymin = mean-ci, ymax = mean+ci), color = "black", 
                width = 0.2, size = 1.25) + 
  scale_color_manual(values = condition.colors) + 
  scale_fill_manual(values = condition.colors) + 
  scale_x_discrete(labels = c("consistent", "typical\nvariation", "atypical\nvariation")) +
  labs(x = "Condition", y = "Accuracy\n(Proportion looking to target object)") + 
  theme_test(base_size = 20) + 
  theme(legend.position = "none")
```

```{r rt}

distracter.initial.trials <- eyetrackingr.data %>%
  filter(time == 0,  fix == "distracter") %>%
  distinct(pid, trial)

rt.data <- eyetrackingr.data %>%
  semi_join(distracter.initial.trials, by = c("pid", "trial")) %>%
  filter(fix == "target" & time > 0) %>%
  group_by(pid, condition, trial, base_word) %>%
  summarize(rt = min(time)) %>%
  ungroup() %>%
  mutate(id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id") %>%
  mutate(id = as.factor(id), 
         condition = relevel(as.factor(condition), ref = "unnaturalv"), 
         z_rt = scale(rt)) %>%
  filter(abs(z_rt) < 3)
  
m <- brm(rt ~ condition + age_c + 
               (1 + condition | id) + 
               (1 + condition + age_c | base_word),
             data = rt.data, 
             family = gaussian(), 
             control = list(adapt_delta = 0.99))

m_null <- update(m, formula = rt ~ age_c + (1 + condition | id) + (1 + condition + age_c| base_word))
loo_compare(loo(m), loo(m_null))

library(bayestestR)

# Posterior summary with 89% credible intervals
posterior_summary(m_brm, prob = 0.89)

# Posterior probabilities of being > 0
p_direction(m_brm)

fixef(m_brm, probs = c(0.055, 0.945))  # for 89% CI



m <- lmer(rt ~ condition + age_c + (1 + condition | id) + (1 | base_word), 
          data = rt.data)

library(lme4)
library(lmerTest)
m <- lmer(prop ~ condition + age_c + (1 + condition | id), 
          data = model.data)

library(glmmTMB)
model.data <- model.data %>%
  mutate(prop_beta = (prop * (nrow(model.data) - 1) + 0.5) / nrow(model.data))

m_beta <- glmmTMB(
  prop_beta ~ condition + 
    (1 | id) + 
    (1 | base_word),
  data = model.data,
  family = beta_family()
)

bypid.bywindow.condition.summary <- data %>%
  filter(!is.na(window)) %>%
  mutate(condition = factor(condition, levels = c("stable", "naturalv", "unnaturalv")), 
         window = factor(window, levels = c("first", "second"), 
                         labels = c("post-naming window 1", 
                                    "post-naming window 2"))) %>%
  group_by(pid, condition, window, age_m) %>%
  summarize(accuracy = mean(fix_num)) %>%
  ungroup()

data %>%
  filter(!is.na(window)) %>%
  mutate(condition = factor(condition, levels = c("stable", "naturalv", "unnaturalv")), 
         window = factor(window, levels = c("first", "second"), 
                         labels = c("post-naming window 1", 
                                    "post-naming window 2"))) %>%
  group_by(pid, condition, window) %>%
  summarize(accuracy = mean(fix_num)) %>%
  ungroup() %>%
  group_by(condition, window) %>%
  summarize(mean = mean(accuracy), 
            sd = sd(accuracy), 
            ci = 1.96*sd/sqrt(n())) %>%
  ggplot(aes(x = condition, y = mean, color = condition, fill = condition)) + 
  facet_grid(.~window) +
  geom_hline(yintercept = 0.5, color = "gray", linetype = "dashed", size = 1.25) + 
  geom_bar(stat = "identity", alpha = 0.6, size = 2) + 
  geom_jitter(data = bypid.bywindow.condition.summary, 
              mapping = aes(x = condition, y = accuracy, color = condition), 
              width = 0.1, size = 3) + 
  geom_errorbar(aes(ymin = mean-ci, ymax = mean+ci), color = "black", 
                width = 0.2, size = 1.25) + 
  scale_color_manual(values = condition.colors) + 
  scale_fill_manual(values = condition.colors) + 
  scale_x_discrete(labels = c("consistent", "typical\nvariation", "atypical\nvariation")) +
  labs(x = "Condition", y = "Accuracy\n(Proportion looking to target object)") + 
  theme_test(base_size = 20) + 
  theme(legend.position = "none")
```
Our primary analysis examines whether children successfully learned the novel word-object mappings in our experiment and compares performance across three different exposure conditions. We used mixed-effects zero-one-inflated beta regression to model the effects of condition (atypical variation vs. consistent vs. typical variation) on accuracy (proportion looking to the target object relative to the total time looking at either the target or distractor object) during the critical analysis window (300-2000ms following target word onset). Age (scaled) was included as an additional fixed effect. We also included the maximal random effects structure (by-child and by-item intercepts and slopes). Beta regression was chosen to model the proportional dependent variable [CITE]. We used a zero-one-inflated model since 0s and 1s are possible values in our data. This particular modeling approach requires a Bayesian framework to interpret. 

```{r cluster-permutation, message=FALSE, warning=FALSE}
library(eyetrackingR)

eyetrackingr.data <- eyetrackingr.data %>%
  mutate(trackloss = NA, 
         aoi = ifelse(target == 1, "target", "distracter"),
         target = ifelse(target == 1, TRUE, FALSE),
         distracter = ifelse(distracter, TRUE, FALSE)) %>% 
  distinct() %>%
  filter(!(pid %in% c("LVar117", "LVar138", "LVar186")))

# Step 1: Prepare the eyetrackingR data object
et_data <- make_eyetrackingr_data(eyetrackingr.data,
                                   participant_column = "pid",
                                   trial_column = "trial",
                                   time_column = "time",  # or "frame" if it's frame-based
                                   trackloss_column = "trackloss",
                                   aoi_columns = c("target", "distracter"),
                                   treat_non_aoi_looks_as_missing = TRUE)

max.time <- data %>%
  summarize(max = mean(second_label) - mean(first_label) + 2000) %>%
  pull(max)

first.window <- 300:2000
second.window <- (mean(data$second_label) - 
                    mean(data$first_label) + 300):(mean(data$second_label) - 
                                                        mean(data$first_label)+2000)

response_window <- subset_by_window(et_data,
                                     window_start_time = min(first.window),
                                     window_end_time = max(second.window), 
                                     rezero = FALSE)

response_window2 <- subset_by_window(et_data,
                                     window_start_time = min(second.window),
                                     window_end_time = max(second.window), 
                                     rezero = FALSE)

response_window1 <- subset_by_window(et_data,
                                     window_start_time = 300,
                                     window_end_time = 2000, 
                                     rezero = FALSE)

response_window_agg_by_sub <- make_time_window_data(response_window, 
                                                    aois='target',
                                              predictor_columns=c('condition'),
                                                    summarize_by = "pid")


plot(response_window_agg_by_sub, predictor_columns="condition", dv = "Prop")

ggplot(response_window_agg_by_sub, aes(x=condition, y = Prop, color = condition)) +
  stat_summary(fun.data = mean_se, geom="pointrange")+
  coord_cartesian(ylim = c(0, 1))+
  labs(y = "Prop. of Target Looks")+
  geom_hline(yintercept = .5, color = "black")+
  scale_color_manual(values = condition.colors) + 
  theme_bw() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

response_time <- make_time_sequence_data(response_window,
                                         time_bin_size = 25, 
                                         predictor_columns = c("condition"),
                                         aois = "target",
                                         summarize_by = "pid")

ggplot(response_time, aes(x = Time, y = Prop))+
  geom_ribbon(aes(xmin = 300, xmax = 2000),
  color = NA, fill = "#9982E1", alpha = 0.1) +
  geom_ribbon(aes(xmin = min(second.window), xmax = max(second.window)),
  color = NA, fill = "#9982E1", alpha = 0.1) +
  stat_smooth(method="loess", span=0.1, se=TRUE, alpha=0.3,
  color = "black", fill = "black") +
  labs(x = "Time from first target label onset (ms)", 
       y = "Accuracy\n(Proportion looking to target object)")+
  geom_hline(yintercept = 0.5, color = "black", linetype="dashed")+
  geom_vline(xintercept = 0, color = "#9982E1", size = 2) +
  geom_vline(xintercept = max.time-2000, color = "#9982E1", size = 2) + 
  scale_x_continuous(limits = c(-2500, 5000), breaks = c(-2500, 0, 2500, 5000)) + 
  coord_cartesian(ylim = c(0,1))+
  theme_classic(base_size = 20) + 
  theme(legend.position = "bottom",
        legend.title = element_blank())

response_time %>%
  ggplot(aes(x = Time, y = Prop, color = condition)) +
  geom_ribbon(aes(xmin = 300, xmax = 2000), color = NA, fill = "gray", alpha = 0.25) + 
  geom_ribbon(aes(xmin = min(second.window), xmax = max(second.window)), 
             color = NA, fill = "gray", alpha = 0.25) + 
  stat_smooth(method="loess", span=0.1, se=TRUE, aes(fill=condition), alpha=0.3) +
  labs(x = "Time from first target label onset (ms)", 
       y = "Accuracy\n(Proportion looking to target object)")+
  geom_hline(yintercept = .5, color = "black", linetype="dashed")+
  geom_vline(xintercept = 0, color = "black")+
  geom_vline(xintercept = max.time-2000, color = "black") + 
  scale_color_manual(values = condition.colors) + 
  scale_fill_manual(values = condition.colors) + 
  coord_cartesian(ylim = c(0,1))+
  theme_classic(base_size = 20) + 
  theme(legend.position = "bottom",
        legend.title = element_blank())

df_timeclust <- make_time_cluster_data(filter(response_time, condition %in% c("stable", "unnaturalv")), 
                                       test= "t.test", paired = TRUE,
                                       predictor_column = "condition", 
                                       threshold = 2)

plot(df_timeclust) +  ylab("T-Statistic") + theme_light() + 
  ggtitle("Both windows: Stable vs. Unnatural variation")
```

## Attention measured by pupil size synchrony

# Discussion

# Conclusion

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# Funding
This research was supported by a XXX awarded to CLW (XXX) and a National Science Foundation Graduate Research Fellowship awarded to KC.

# Acknowledgements
We are grateful to the participating families as well as the members of the Princeton Baby Lab for their feedback on this work.
