---
title: "Supplemental Materials for: English-hearing toddlers learn words from consistent and variable labels"
shorttitle: "SM: Learning from consistent and variable labels"
author:
  - name: Kennedy Casey
    corresponding: true
    orcid: 0000-0001-5326-3080
    email: kcasey@princeton.edu
    affiliations:
      - name: Princeton University
        department: Department of Psychology
        address: 40 Woodlands Way
        city: Princeton
        region: NJ
        country: USA
        postal-code: 08544
  - name: Casey Lew-Williams
    orcid: 0000-0002-8781-4458
    email: caseylw@princeton.edu
    affiliations:
      - name: Princeton University
        department: Department of Psychology
        address: 40 Woodlands Way
        city: Princeton
        region: NJ
        country: USA
        postal-code: 08544
author-note:
  disclosures:
    study-registration: "The study design and analysis plan was pre-registered ([https://osf.io/uqz6d](https://osf.io/uqz6d))"
    data-sharing: "Anonymized data are available at LINK"
    conflict-of-interest: "The authors declare no conflicts of interest"
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "This document is a template."
keywords: [word learning, child-directed speech, wordplay, eye-tracking]
bibliography: LexVar.bib
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
execute: 
  echo: false
  warning: false
  message: false
  error: false
  eval: true

---
```{r setup}
library(tidyverse)
library(readxl)
library(english)
library(tools)
library(flextable)
library(brms)
library(loo)
library(emmeans)

# reprocess raw eye-tracking data? set to TRUE or FALSE
reprocess = FALSE

# reprocess pupillometry data? set to TRUE or FALSE
reprocess.pupil = FALSE 

if (reprocess) {
  source("analysis/process.R")
}

if (reprocess.pupil) {
  source("analysis/pupillometry.R")
}

# set condition colors
condition.colors <- c("stable" = "#ef5039", 
                      "unnaturalv" = "#04a48c", 
                      "naturalv" = "#3f5eaa")
```

# Method
This study was pre-registered ([https://osf.io/uqz6d](https://osf.io/uqz6d)). All stimuli, anonymized data, and analysis scripts are publicly available at LINK.

```{r, echo=FALSE}
log <- read_xlsx("data/metadata/ANON-ptcp-log.xlsx")
final.sample <- filter(log, usable_checked == "y") 
usable <- final.sample %>% nrow()
unusable <- filter(log, usable_checked == "n") %>% nrow() - 1 # subtracting since 1 child was accidentally double recruited
unusable.cal <- filter(log, str_detect(exclusion_reason, "calibration|refused")) %>% nrow()
unusable.data <- filter(log, str_detect(exclusion_reason, "trial")) %>% nrow()
unusable.inter <- filter(log, str_detect(exclusion_reason, "interruption")) %>% nrow()
unusable.lang <- filter(log, str_detect(exclusion_reason, "language")) %>% nrow()

ethnicity <- final.sample %>%
  mutate(ethnicity = case_when(
    ethnicity %in% c("black", "asian", "not reported") ~ str_to_sentence(ethnicity), 
    ethnicity == "white" ~ "Non-Latinx white", 
    ethnicity == "white/hispanic" ~ "Latinx white", 
    ethnicity %in% c("white/black", "asian/white", "mixed") ~ "Mixed race"
  )) %>%
  group_by(ethnicity) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = round(n/sum(n)*100, 2)) %>%
  arrange(desc(prop))

sex <- final.sample %>%
  group_by(sex) %>%
  summarize(n = n())

ages <- final.sample %>%
  transmute(id = str_extract(ptcp_id, "[0-9][0-9][0-9]"), 
            age_c = datawizard::standardize(age))
```

## Participants
Participants (_N_ = `r usable` total, `r filter(sex, sex == "F")$n` female, `r filter(sex, sex == "M")$n` male,), aged 18 to 36 months (*M*~age~ = `r format(round(mean(final.sample$age), 2), nsmall = 2)`), were full-term, monolingual English-hearing children (>80% reported exposure) with no known hearing, vision, language, and/or developmental delays. Caregivers self-reported race/ethnicity information for their children in an open-ended format when they initially signed up for our volunteer database. `r sub("^([a-z])", "\\U\\1", as.character(english(ethnicity[3,]$n)), perl = TRUE)` caregivers  (`r ethnicity[3,]$prop`% of the final sample) chose not to report this information. Participants were primarily `r ethnicity[1,]$ethnicity` (_N_ = `r ethnicity[1,]$n`, `r ethnicity[1,]$prop`%) with some representation of other races/ethnicities: Mixed race (_N_ = `r ethnicity[2,]$n`, `r ethnicity[2,]$prop`%), Asian (_N_ = `r ethnicity[4,]$n`, `r ethnicity[4,]$prop`%), Black (_N_ = `r ethnicity[5,]$n`, `r ethnicity[5,]$prop`%), `r ethnicity[6,]$ethnicity` (_N_ = `r ethnicity[6,]$n`, `r ethnicity[6,]$prop`%).

The final sample size was pre-registered and determined using an a priori power analysis conducted in G*Power [@faul2009statistical] and informed by prior eye-tracking studies with a similar age range [@nencheva2021moment; @quam2023protracted]. `r sub("^([a-z])", "\\U\\1", as.character(english(unusable)), perl = TRUE)` additional children participated but were excluded from the final analyses due to refusal to wear the EyeLink target sticker or failure to complete calibration (_N_ = `r unusable.cal`), significant interruption by a family member (_N_ = `r unusable.inter`), contribution of too few usable trials based on pre-registered criteria due to fussiness or inattention (_N_ = `r unusable.data`), or failure to meet the language requirement (_N_ = `r unusable.lang`).

All experimental protocols, including procedures for obtaining informed consent, were approved by the Princeton University Institutional Review Board (Approval: 0000007117, Language Learning: Sounds, Words, and Grammar).

## Stimuli
```{r stimuli}
library(tuneR)
training.audio <- "stimuli/training_audio/"
testing.audio <- "stimuli/testing_audio/"

training.wavs <- list.files(path = training.audio, pattern = "\\.wav$", full.names = TRUE)
testing.wavs <- list.files(path = testing.audio, pattern = "\\.wav$", full.names = TRUE)

training.durs <- tibble(file = training.wavs) %>%
  mutate(
    name = basename(file),
    wav = map(file, readWave),
    dur = round(map_dbl(wav, ~ length(.x@left) / .x@samp.rate) * 1000, 0)
  ) %>%
  summarize(mean = round(mean(dur), 0),
            min = min(dur), 
            max = max(dur))

test.word.durs <- tibble(base_word = "bog", window = "first", dur = 615) %>%
  add_row(base_word = "bog", window = "second", dur = 613) %>%
  add_row(base_word = "mob", window = "first", dur = 621) %>%
  add_row(base_word = "mob", window = "second", dur = 621) %>%
  add_row(base_word = "kai", window = "first", dur = 622) %>%
  add_row(base_word = "kai", window = "second", dur = 628) %>%
  add_row(base_word = "tan", window = "first", dur = 625) %>%
  add_row(base_word = "tan", window = "second", dur = 628) %>%
  add_row(base_word = "voo", window = "first", dur = 630) %>%
  add_row(base_word = "voo", window = "second", dur = 626) %>%
  add_row(base_word = "woi", window = "first", dur = 627) %>%
  add_row(base_word = "voo", window = "second", dur = 634) %>%
  summarize(mean = round(mean(dur), 0),
            min = min(dur), 
            max = max(dur))
  
testing.dur <- tibble(file = testing.wavs) %>%
  mutate(
    name = basename(file),
    wav = map(file, readWave),
    dur = round(map_dbl(wav, ~ length(.x@left) / .x@samp.rate) * 1000, 0)
  ) %>%
  select(dur) %>%
  distinct() %>%
  pull(dur) - 4000 # accounting for 2 x 2000ms silent periods before and after trial
```
Auditory stimuli were recorded by female native English speaker using naturalistic child-directed prosody in a sound-attenuated room. Each target word (see Table 1) appeared in four different sentence frames for exposure trials ("Look, _____! That's a _____. Do you see the ______? Yes, that's a ______.") and in two different sentence frames for test trials ("Can you find the _____? See the _____?"). The exact same sentence frame recordings were used across all target words. Trials were created in Praat [@praat] by concatenating each target word onto the final position of each sentence frame. To avoid co-articulation artifacts in trials with articles preceding the target word, the target words were spliced along with the article. Exposure trials were `r format(training.durs$mean, scientific = FALSE, big.mark = ",")`ms long, on average (range = `r format(training.durs$min, scientific = FALSE, big.mark = ",")`--`r format(training.durs$max, scientific = FALSE, big.mark = ",")`ms), including 2000ms silent periods between exposure sentences. Test trials were all exactly `r testing.dur`ms long, with target word durations ranging from `r test.word.durs$min` to `r test.word.durs$max`ms (_M_ = `r test.word.durs$mean`ms) and including 2000ms silence between target sentences. All auditory stimuli were normed to a standard mean intensity of 65dB. Minor acoustic edits were made on some sentences to ensure that all trials sounded naturally produced and had comparable loudness and pitch.

```{r stimuli-table}
tribble(
  ~Standard, ~`Reduplicated (Typical)`, ~`Reduplicated (Atypical)`, ~`Diminutivized (Typical)`, ~`Diminutivized (Atypical)`,
  "tanzer\n/tæn.zɚ/", "tan-tan\n/tæn.tæn/", "zer-zer\n/zɚ.zɚ/", "tanny\n/tæn.i/", "y-zer\n/i.zɚ/",
  "bogdum\n/bɑg.dʌm/", "bog-bog\n/bɑg.bɑg/", "dum-dum\n/dʌm.dʌm/", "boggy\n/bɑg.i/", "y-dum\n/i.dʌm/",
  "kaisol\n/kaɪ.sɔl/", "kai-kai\n/kaɪ.kaɪ/", "sol-sol\n/sɔl.sɔl/", "kaisy\n/kaɪ.si/", "y-sol\n/i.sɔl/",
  "mobet\n/moʊ.bɛt/", "mo-mo\n/moʊ.moʊ/", "bet-bet\n/bɛt.bɛt/", "moby\n/moʊ.bi/", "y-bet\n/i.bɛt/",
  "woifer\n/wɔɪ.fɚ/", "woi-woi\n/wɔɪ.wɔɪ/", "fer-fer\n/fɚ.fɚ/", "woify\n/wɔɪ.fi/", "y-fer\n/i.fɚ/",
  "voopin\n/vu.pɪn/", "voo-voo\n/vu.vu/", "pin-pin\n/pɪn.pɪn/", "voopy\n/vu.pi/", "y-pin\n/i.pɪn/"
) %>% flextable() %>%
  set_caption("Table 1. Target word forms.") %>%
  set_table_properties(width = 1, layout = "autofit") %>%
  fontsize(size = 10, part = "all") %>%
  align(align = "left", part = "all")
```

The visual stimuli consisted of six isoluminant objects with 12 unique colors (two per object) that were modeled after @nencheva2021moment's stimuli and hand-drawn using Procreate software [@procreate].

## Procedure
Toddlers sat on their caregiver's lap, approximately 60cm away from a 17-inch monitor equipped with an EyeLink 1000 Plus eye-tracker, in a dimly-lit testing room. Caregivers wore opague sunglasses to prevent them from influencing their children's behavior. Participants wore a small EyeLink target sticker on their forehead. We used a 3-point calibration procedure along the horizontal dimension (x-axis) of the screen. 

The experiment consisted of three blocks (1 per condition). Each block consisted of two phases: an exposure phase, during which participants learned novel word-object pairings, and a test phase, during which toddlers were immediately tested on their recognition of the word-object mappings. An isoluminant, non-linguistic attention-getting video was shown after every block.

During the exposure phase (4 trials per block), children saw a single target object appear at the center of the screen. Before the onset of each trial, children saw a blank gray screen. Each trial started after a short, variable-length ITI (min: 250ms, max: 750ms, mean: 500ms) or when the participant looked at the screen. Next, the target object appeared silently on the screen for 1500ms or until the participant looked at the screen. Then, the exposure audio played (see stimulus details above). After the offset of the audio, the target object remained on the screen for an additional 1500ms.

During the test phase (4 trials per block), toddlers saw the two objects from the preceding exposure phase side-by-side on the screen. Each trial started with an ITI (same structure as training phase), followed by a 2000ms silent period with both objects on the screen. Then, the target audio played (see stimulus details above). After the offset of the audio, the side-by-side objects remained on the screen for an additional 2000ms.

## Conditions
Participants learned words in three conditions (within-subjects manipulation). In all conditions, participants learned the names of two different novel objects across four alternating exposure trials (2 per word). In the _consistent_ condition, participants heard the standard form of the target word repeated 5x per trial (e.g., "Look, tanzer! That's a tanzer...tanzer. Do you see the tanzer? Yes, that's a tanzer."). In the _typical variation_ condition, participants heard the standard form of the target word repeated 2x per trial in addition to one reduplicated form and one diminutivized form per trial (e.g., "Look, tanzer! That's a tan-tan. Do you see the tanny? Yes, that's a tanzer."). Critically, the number of exposures to the first syllable of the target word (i.e., "tan") is equivalent across these two conditions. In the _atypical variation_ condition, participants heard the standard form of the target word repeated 2x per trial in addition to 1x exposure each of reversed versions of the typical reduplicated and diminutivized forms (e.g., "Look, tanzer! That's a zer-zer. Do you see the y-zer? Yes, that's a tanzer."). A full list of exposed word forms can be found in Table 1. 

Participants were assigned to one of 18 pseudorandomized orders such that the order of presentation of blocks and the assignment of target labels to target objects was counterbalanced across conditions.

```{r usable-stats}
path <- "data/processed/usable/"
usable <- tibble()
for (i in list.files(path)) {
  usable <- bind_rows(usable, 
                      read_csv(paste0(path, i), show_col_types = FALSE))
}

n.conditions <- usable %>%
  group_by(id) %>%
  summarize(n = length(unique(condition))) %>%
  ungroup() %>%
  summarize(mean = round(mean(n), 2),
            median = round(median(n), 2),
            min = min(n),
            max = max(n))

n.trials <- usable %>%
  group_by(id) %>%
  summarize(n = length(unique(trial))) %>%
  ungroup() %>%
  summarize(mean = round(mean(n), 2),
            median = round(median(n), 2),
            min = min(n),
            max = max(n))
```
## Data pre-processing
Gaze behavior was used to determine toddlers' recognition of novel word-object pairings during two critical windows of analysis (300ms-2000ms following each target word onset; @fernald2008looking; @swingley2000spoken). We calculated participants' proportion looking to target objects during test trials as the measure of word-recognition accuracy. For any given test trial to be included in analyses, children must have looked to the target and/or distractor object at least once prior to the onset of the label and for at least 1/3 of each critical post-labeling window. In order for a block to be included, children must have contributed usable data for at least 2 of 4 test trials. Finally, in order for a participant to be included, they must have contributed at least 2 of 3 blocks. On average, participants contributed `r n.trials$mean` trials (median = `r n.trials$median`, range = `r n.trials$min`--`r n.trials$max`) across `r n.conditions$mean` conditions (median = `r n.conditions$median`).

Following @nencheva2021moment, we first converted toddlers' raw pupil size values during exposure sentences (automatically recorded by the EyeLink eye-tracker) to percent change the average pupil size for that trial. We removed potential artifacts due to partial blinks by excluding any 0.05-s intervals with more than 15% change in pupil size [@merritt1994artifact]. We used Stineman interpolation to fill in gaps shorter than 100 ms using Stineman interpolation [@stinepack]. Finally, we excluded segments where more than 50% of the data were missing.

# Results
## Word recognition measured by gaze behavior
```{r test-prep}
rewrite = FALSE
path <- "data/processed/"

if (rewrite) {
  test.props <- tibble()
  test.props.bywindow <- tibble()
  frames <- tibble()
  frames.cbp <- tibble()
  
  included.files <- list.files(path, "-test.csv")
  
  for (i in included.files) {
    
    curr.file <- read_csv(paste0(path, i))
    
    id <- str_extract(unique(curr.file$pid), "[0-9][0-9][0-9]")
    
    status <- filter(log, str_detect(ptcp_id, id))$usable_checked
    
    if (status != "n" & !is.na(status)) {
      # read in usable trials from pre-processing stage
      usable.trials <- read_csv(paste0("data/processed/usable/", id, ".csv")) %>%
        pull(trial)
      
      if (length(usable.trials) > 0) {
        
      curr.file.cbp <- curr.file %>% 
        select(pid, age_m, order, trial, x, pupil, base_word:window) %>%
        filter(trial %in% usable.trials) %>%
        group_by(pid, trial, condition, fix, base_word, window) %>%
        filter(length(unique(frame)) >= (2000-300)/3) %>%
        ungroup()
        
      curr.file <- curr.file %>%
        select(pid, age_m, order, trial, x, pupil, base_word:window) %>%
        filter(trial %in% usable.trials & !is.na(window)) %>%
        group_by(pid, trial, condition, fix, base_word, window) %>%
        filter(length(unique(frame)) >= (2000-300)/3) %>%
        ungroup()

      curr.test.props <- curr.file %>%
        group_by(pid, trial, condition, fix, base_word) %>%
        summarize(total = n()) %>%
        pivot_wider(names_from = "fix", values_from = "total",
                    id_cols = c(pid, trial, condition, base_word))

      curr.test.props.bywindow <- curr.file %>%
        group_by(pid, trial, condition, fix, base_word, window) %>%
        summarize(total = n()) %>%
        pivot_wider(names_from = "fix", values_from = "total",
                    id_cols = c(pid, trial, condition, base_word, window))

      if ("target" %in% colnames(curr.test.props)) {
        curr.test.props <- curr.test.props %>%
          mutate(target = ifelse(is.na(target), 0, target))

      } else {
        curr.test.props <- curr.test.props %>%
          mutate(target = 0)
      }

      if ("distracter" %in% colnames(curr.test.props)) {
        curr.test.props <- curr.test.props %>%
          mutate(distracter = ifelse(is.na(distracter), 0, distracter))
      } else {
        curr.test.props <- curr.test.props %>%
          mutate(distracter = 0)
      }

      curr.test.props <- curr.test.props %>%
        mutate(total = target + distracter,
               prop = target/total)

      test.props <- bind_rows(test.props, curr.test.props)

      if ("target" %in% colnames(curr.test.props.bywindow)) {
        curr.test.props.bywindow <- curr.test.props.bywindow %>%
          mutate(target = ifelse(is.na(target), 0, target))

      } else {
        curr.test.props.bywindow <- curr.test.props.bywindow %>%
          mutate(target = 0)
      }

      if ("distracter" %in% colnames(curr.test.props.bywindow)) {
        curr.test.props.bywindow <- curr.test.props.bywindow %>%
          mutate(distracter = ifelse(is.na(distracter), 0, distracter))
      } else {
        curr.test.props.bywindow <- curr.test.props.bywindow %>%
          mutate(distracter = 0)
      }

      curr.test.props.bywindow <- curr.test.props.bywindow %>%
        mutate(total = target + distracter,
               prop = target/total)

      test.props.bywindow <- bind_rows(test.props.bywindow, curr.test.props.bywindow)

      curr.frames <- curr.file %>%
        select(pid, age_m, order, trial, x, pupil, base_word:window) %>%
        filter(trial %in% usable.trials) %>%
        mutate(fix_num = case_when(
          fix == "distracter" ~ 0,
          fix == "target" ~ 1,
          TRUE ~ NA
        ))
      
      curr.frames.cbp <- curr.file.cbp %>%
        select(pid, age_m, order, trial, x, pupil, base_word:window) %>%
        filter(trial %in% usable.trials) %>%
        mutate(fix_num = case_when(
          fix == "distracter" ~ 0, 
          fix == "target" ~ 1, 
          TRUE ~ NA
        ))
      
      frames <- bind_rows(frames, curr.frames)
      frames.cbp <- bind_rows(frames.cbp, curr.frames.cbp)
      
      }
    }
  }
  
  # export proportion looking data
  write_csv(test.props, "data/test-props.csv")
  write_csv(test.props.bywindow, "data/test-props-bywindow.csv")
  
  # export full timecourse data
    data <- frames %>%
      mutate(time = frame,
             condition = factor(condition, levels = c("unnaturalv", "stable", "naturalv"))) %>%
      select(pid, age_m, order, trial, condition, base_word, pupil, time, fix_num, window,
             first_label, second_label) %>%
      group_by(pid) %>%
      mutate(block_order = list(unique(condition))) %>%
      ungroup() %>%
      mutate(condition_order = map2_int(block_order, condition, ~which(.x == .y)))
    write_csv(data, "data/prepped-data.csv")
    
    # export eyetrackingR version of data
    eyetrackingr.data <- frames.cbp %>%
      mutate(time = frame, 
             condition = factor(condition, levels = c("unnaturalv", "stable", "naturalv"))) %>%
      mutate(target = ifelse(fix_num == 1, 1, 0),
             distracter = ifelse(fix_num == 0, 1, 0)) %>%
      select(-fix_num)
    write_csv(eyetrackingr.data, "data/eyetrackingr-data.csv")
} else {
  data <- read_csv("data/prepped-data.csv")
  test.props <- read_csv("data/test-props.csv")
  test.props.bywindow <- read_csv("data/test-props-bywindow.csv")
  eyetrackingr.data <- read_csv("data/eyetrackingr-data.csv")
}
```

```{r test-bayes, results = "hide"}
model.data <- test.props %>%
  mutate(id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id") %>%
  mutate(condition = relevel(as.factor(condition), ref = "unnaturalv"))

model.data.bywindow <- test.props.bywindow %>%
  mutate(id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id") %>%
  mutate(condition = relevel(as.factor(condition), ref = "unnaturalv"))

rerun = FALSE
if (rerun) {
  m.age <- brm(
    bf(prop ~ condition + age_c + (1 + condition | id) + (1 + condition + age_c | base_word),
     phi ~ condition + age_c,
     zoi ~ condition + age_c,
     coi ~ condition + age_c),
  data = model.data,
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  save_pars = save_pars(all = TRUE),
  iter = 2000,
  warmup = 1000,
  seed = 20250803,
  control = list(adapt_delta = 0.999)
  )
  saveRDS(m.age, file = "models/m-age.rds")
  
  m.age.int <- brm(
  bf(prop ~ condition * age_c + (1 + condition | id) + (1 + condition * age_c | base_word),
     phi ~ condition * age_c,
     zoi ~ condition * age_c,
     coi ~ condition * age_c),
  data = model.data,
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  save_pars = save_pars(all = TRUE),
  iter = 2000,
  warmup = 1000,
  seed = 20250803,
  control = list(adapt_delta = 0.999)
  )
  saveRDS(m.age.int, file = "models/m-age-int.rds")
  
  m.window1 <- brm(
  bf(prop ~ condition + age_c + (1 + condition | id) + (1 + condition + age_c | base_word),
     phi ~ condition + age_c,
     zoi ~ condition + age_c,
     coi ~ condition + age_c),
  data = filter(model.data.bywindow, window == "first"),
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 20250803,
  control = list(adapt_delta = 0.999)
)
  saveRDS(m.window1, file = "models/m-window1.rds")

  m.window2 <- brm(
  bf(prop ~ condition + age_c + (1 + condition | id) + (1 + condition + age_c | base_word),
     phi ~ condition + age_c,
     zoi ~ condition + age_c,
     coi ~ condition + age_c),
  data = filter(model.data.bywindow, window == "second"),
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 2000,
  warmup = 1000,
  seed = 20250803,
  control = list(adapt_delta = 0.999)
)
  saveRDS(m.window2, file = "models/m-window2.rds")
  
  # compare models (pre-registered plan)
  loo1 <- loo(m.age, moment_match = TRUE, reloo = TRUE)
  loo2 <- loo(m.age.int, moment_match = TRUE, reloo = TRUE)
  comp <- loo_compare(loo1, loo2) %>%
    as_tibble() %>%
    tibble::rownames_to_column(var = "model") %>%
    write_csv("models/compare.csv")
} else {
  m.age <- readRDS("models/m-age.rds")
  m.age.int <- readRDS("models/m-age-int.rds")
  m.window1 <- readRDS("models/m-window1.rds")
  m.window2 <- readRDS("models/m-window2.rds")
  comp <- read_csv("models/compare.csv")
}

first.window <- 300:2000
second.window <- (mean(data$second_label) - 
                    mean(data$first_label) + 300):(mean(data$second_label) - 
                                                        mean(data$first_label)+2000)

library(tidybayes)
# summarize posterior
m.age_summary <- as_draws_df(m.age) %>%
    pivot_longer(cols = starts_with("b_"), 
                 names_to = "parameter", 
                 values_to = "value") %>%
  group_by(parameter) %>%
  mean_qi(.width = 0.89) %>%
  select(parameter, starts_with("value"))

m.window1_summary <- as_draws_df(m.window1) %>%
    pivot_longer(cols = starts_with("b_"), 
                 names_to = "parameter", 
                 values_to = "value") %>%
  group_by(parameter) %>%
  mean_qi(.width = 0.89) %>%
  select(parameter, starts_with("value"))

m.window2_summary <- as_draws_df(m.window2) %>%
    pivot_longer(cols = starts_with("b_"), 
                 names_to = "parameter", 
                 values_to = "value") %>%
  group_by(parameter) %>%
  mean_qi(.width = 0.89) %>%
  select(parameter, starts_with("value"))

conditional_effects(m.age)

# generate posterior predictions per condition
newdata <- data.frame(
  condition = factor(c("unnaturalv", "stable", "naturalv"), levels = levels(model.data$condition)), 
  age_c = 0
)

# get fitted estimates of the mean (mu parameter of the beta)
preds <- posterior_epred(m.window2, newdata = newdata, re_formula = NA)

# calculate means and 89% credible intervals
pred_summary <- as.data.frame(preds) %>%
  setNames(levels(newdata$condition)) %>%
  pivot_longer(everything(), names_to = "condition", values_to = "mu") %>%
  group_by(condition) %>%
  summarize(
    mean = mean(mu),
    ci_lower = quantile(mu, 0.055),
    ci_upper = quantile(mu, 0.945)
  )

print(pred_summary)
```
```{r test-bayes-reporting}
library(tidybayes)
library(posterior)
draws <- as_draws_df(m.age)
summary <- summarize_draws(draws)
```

We fit a Bayesian zero-one inflated beta regression model [@liu2018review] to examine the effects of input condition and age on children's accuracy of novel word recognition (i.e., proportion looking to the target object), using the `brms` package [@brms]. The model included fixed effects for condition (atypical variation vs. typical variation vs. consistent) and mean-centered age, along with the maximal random effects structure (by-participant and by-item random slopes and intercepts). All R-hat values were 1.00, indicating good convergence. Effective sample sizes (Bulk ESS: `r round(min(summary$ess_bulk, na.rm = TRUE), 0)`--`r round(max(summary$ess_bulk, na.rm = TRUE), 0)`; Tail ESS: `r round(min(summary$ess_tail, na.rm = TRUE), 0)`--`r round(max(summary$ess_tail, na.rm = TRUE), 0)`) were sufficiently large, indicating stable and reliable posterior estimates.

Following our pre-registered analysis plan, we tested whether including an interaction between condition and age improved model fit---we compared the model with the interaction term to the simpler additive model using approximate leave-one-out cross-validation [@loo]. This analysis suggested no reliable difference in predictive accuracy, providing no strong evidence that the interaction term improved model fit (ΔELPD = `r format(round(filter(comp, model == 2)$elpd_diff, 2), nsmall = 2)`, ΔSE = `r format(round(filter(comp, model == 2)$se_diff, 2), nsmall = 2)`).

The beta component of the model, which estimates the effects on the mean outcome for non-boundary outcomes (i.e., values between but not including 0 or 1) showed small positive effects for both the natural variation condition (β = `r format(round(filter(m.age_summary, parameter == "b_conditionnaturalv")$value, 2), nsmall = 2)`, 89% CrI: [`r format(round(filter(m.age_summary, parameter == "b_conditionnaturalv")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.age_summary, parameter == "b_conditionnaturalv")$value.upper, 2), nsmall = 2)`]) and the consistent condition (β = `r format(round(filter(m.age_summary, parameter == "b_conditionstable")$value, 2), nsmall = 2)`, 89% CrI: [`r format(round(filter(m.age_summary, parameter == "b_conditionstable")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.age_summary, parameter == "b_conditionstable")$value.upper, 2), nsmall = 2)`]), relative to the unnatural variation condition. Given that the credible intervals include zero, the results do not provide strong evidence for a condition effect. The effect of centered age was negligible (β = `r format(round(filter(m.age_summary, parameter == "b_age_c")$value, 2), nsmall = 2)`, 89% CrI: [`r format(round(filter(m.age_summary, parameter == "b_age_c")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.age_summary, parameter == "b_age_c")$value.upper, 2), nsmall = 2)`])

The conditional one-inflation (coi) component, which predicts---among zero/one responses---the probability of being 1 (vs. 0), also showed similar trends for condition. The effect of the typical variation condition relative to atypical variation was weakly positive, and the credible intervals include zero (β = `r format(round(filter(m.age_summary, parameter == "b_coi_conditionnaturalv")$value, 2), nsmall = 2)`, 89% CrI: [`r format(round(filter(m.age_summary, parameter == "b_coi_conditionnaturalv")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.age_summary, parameter == "b_coi_conditionnaturalv")$value.upper, 2), nsmall = 2)`]). The effect of the stable condition relative to the atypical variation condition was stronger but still uncertain given the credible interval overlap with zero. (β = `r format(round(filter(m.age_summary, parameter == "b_coi_conditionstable")$value, 2), nsmall = 2)`, 89% CrI: [`r format(round(filter(m.age_summary, parameter == "b_coi_conditionstable")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.age_summary, parameter == "b_coi_conditionstable")$value.upper, 2), nsmall = 2)`]). The effect of age was clearly positive (β = `r format(round(filter(m.age_summary, parameter == "b_coi_age_c")$value, 2), nsmall = 2)`, 89% CrI: [`r format(round(filter(m.age_summary, parameter == "b_coi_age_c")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.age_summary, parameter == "b_coi_age_c")$value.upper, 2), nsmall = 2)`]), indicating that older children were more likely to perform at ceiling (i.e., demonstrate 100% looking to the target object during the critical analysis window).

```{r extract-summary, echo=FALSE, execute = FALSE}
library(brms)
library(dplyr)
library(tibble)

# Extract posterior summaries
post_sum <- posterior_summary(m.window1, prob = 0.89)

# Pull out fixed effects
natv_est <- post_sum["b_conditionnaturalv", "Estimate"]
natv_sd <- post_sum["b_conditionnaturalv", "Est.Error"]
natv_lb <- natv_est - qnorm(0.945) * natv_sd
natv_ub <- natv_est + qnorm(0.945) * natv_sd

stable_est <- post_sum["b_conditionstable", "Estimate"]
stable_sd <- post_sum["b_conditionstable", "Est.Error"]
stable_lb <- stable_est - qnorm(0.945) * stable_sd
stable_ub <- stable_est + qnorm(0.945) * stable_sd

age_est <- post_sum["b_age_c", "Estimate"]
age_sd <- post_sum["b_age_c", "Est.Error"]
age_lb <- age_est - qnorm(0.945) * age_sd
age_ub <- age_est + qnorm(0.945) * age_sd

# Posterior probabilities
draws <- as_draws_df(m.window1)
prob_natv <- mean(draws$b_conditionnaturalv > 0)
prob_stable <- mean(draws$b_conditionstable > 0)
prob_age <- mean(draws$b_age_c > 0)
```

Condition effects were weakly positive but uncertain. For the typical variation condition, the estimated effect was `r round(natv_est, 3)`, SD = `r round(natv_sd, 3)`, with a 89% credible interval from `r round(natv_lb, 3)` to `r round(natv_ub, 3)`. For the consistent condition, the estimate was `r round(stable_est, 3)`, SD = `r round(stable_sd, 3)`, 89% CI [`r round(stable_lb, 3)`, r `round(stable_ub, 3)`]. The effect of age was small (`r round(age_est, 3)`, SD = `r round(age_sd, 3)`), 89% CI [`r round(age_lb, 3)`, `r round(age_ub, 3)`].

Posterior probabilities that each effect was greater than zero were: `r round(prob_natv, 2)` for the effect of the typical variation condition (relative to atypical variation condition), `r round(prob_stable, 2)` for the effect of the consistent condition (relative to the atypical variation condition), and `r round(prob_age, 2)` for the effect of age. These results suggest limited evidence that either input condition or age significantly affected performance, though the posterior distributions for all three predictors leaned slightly positive.

```{r fig1}
#| label: fig1
#| fig-cap: "Figure 1: Timecourse of participants' looks during test trials with shaded standard error regions. The gray dashed line at 0.5 represents chance looking. The black dashed lines indicate the critical analysis window (300-2000ms) following target word onset."
#| width: 12
#| height: 4

eyetrackingr.data %>% 
  mutate(accuracy = ifelse(fix == "target", 1, 0)) %>%
  filter(time >= -1500 & time <= 2500) %>%
  group_by(pid, time, condition, trial) %>%
  summarize(accuracy = mean(accuracy, na.rm = TRUE)) %>% 
  group_by(time, condition) %>% # aggregate across participants
  summarize(
    N = sum(!is.na(accuracy)),
    SD = sd(accuracy, na.rm = TRUE),
    SE = SD/sqrt(N),
    accuracy = mean(accuracy, na.rm = TRUE),
    lower = accuracy-SE,
    upper = accuracy+SE
  ) %>%
  ggplot(aes(x=time, y=accuracy, color = condition)) + 
  scale_fill_manual(values = condition.colors, labels = c("Typical variation", "Consistent", "Atypical variation")) +
  scale_color_manual(values = condition.colors, labels = c("Typical variation", "Consistent", "Atypical variation")) +
  geom_hline(aes(yintercept=0.5),linetype="dashed",color='gray') + 
  geom_smooth(aes(ymin=lower, ymax=upper, fill = condition), stat="identity") +
  geom_vline(xintercept = 300, linetype = "dashed", color = "black") +
  geom_vline(xintercept = 2000, linetype = "dashed", color = "black") +
  geom_line() +
  theme_test(base_size=15) +
  coord_cartesian(xlim=c(-500,2500), ylim= c(-.01,1.01),expand=F) +
  scale_x_continuous(breaks=seq(from=-500,to=2500,by=200)) +
  labs(x="Time from target word onset onset (ms)",
       y="Accuracy\n(Proportion looking to target object)", 
       color = "Condition", 
       fill = "Condition") +
  theme(axis.title = element_text(face = "bold"), 
        legend.title = element_text(face = "bold"), 
        legend.position = "bottom",
        axis.text.x = element_text(size = 8))
```

### Window effects

In the early time window, there were no strong condition effects. The posterior estimate for the typical variation condition relative to atypical variation was β = `r format(round(filter(m.window1_summary, parameter == "b_conditionnaturalv")$value, 2), nsmall = 2)` (89% CrI: \[`r format(round(filter(m.window1_summary, parameter == "b_conditionnaturalv")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.window1_summary, parameter == "b_conditionnaturalv")$value.upper, 2), nsmall = 2)`\]). Similarly, the consistent condition estimate was β = `r format(round(filter(m.window1_summary, parameter == "b_conditionstable")$value, 2), nsmall = 2)` (89% CrI: \[`r format(round(filter(m.window1_summary, parameter == "b_conditionstable")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.window1_summary, parameter == "b_conditionstable")$value.upper, 2), nsmall = 2)`\]). These effects were small, and their credible intervals included zero.

In the later window, condition effects were descriptively stronger. The estimate for the typical variation condition relative to atypical variation was β = `r format(round(filter(m.window2_summary, parameter == "b_conditionnaturalv")$value, 2), nsmall = 2)` (89% CrI: \[`r format(round(filter(m.window2_summary, parameter == "b_conditionnaturalv")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.window2_summary, parameter == "b_conditionnaturalv")$value.upper, 2), nsmall = 2)`\]), and for the consistent condition: β = `r format(round(filter(m.window2_summary, parameter == "b_conditionstable")$value, 2), nsmall = 2)` (89% CrI: \[`r format(round(filter(m.window2_summary, parameter == "b_conditionstable")$value.lower, 2), nsmall = 2)`, `r format(round(filter(m.window2_summary, parameter == "b_conditionstable")$value.upper, 2), nsmall = 2)`\]). Though credible intervals still overlapped zero, the magnitude and direction of the effects in this window were more consistent with our hypotheses.



```{r rt, results = "hide"}
distracter.initial.trials <- eyetrackingr.data %>%
  filter(time == 0,  fix == "distracter") %>%
  distinct(pid, trial)

rt.data <- eyetrackingr.data %>%
  semi_join(distracter.initial.trials, by = c("pid", "trial")) %>%
  filter(fix == "target" & time > 0) %>%
  group_by(pid, condition, trial, base_word) %>%
  summarize(rt = min(time)) %>%
  ungroup() %>%
  mutate(id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id") %>%
  mutate(id = as.factor(id), 
         condition = relevel(as.factor(condition), ref = "unnaturalv"), 
         z_rt = scale(rt)) %>%
  filter(abs(z_rt) < 3)

if (rerun) {
  m.rt <- brm(rt ~ condition + age_c + 
               (1 + condition | id) + 
               (1 + condition + age_c | base_word),
             data = rt.data, 
             family = gaussian(), 
             control = list(adapt_delta = 0.99))
  saveRDS(m.rt, file = "models/m-rt.rds")
} else {
  m.rt <- readRDS(file = "models/m-rt.rds")
}

library(bayestestR)

# Posterior summary with 89% credible intervals
posterior_summary(m.rt, prob = 0.89)

# Posterior probabilities of being > 0
p_direction(m.rt)

fixef(m.rt, probs = c(0.055, 0.945))  # for 89% CI
```
Our primary analysis examines whether children successfully learned the novel word-object mappings in our experiment and compares performance across three different exposure conditions. We used mixed-effects zero-one-inflated beta regression to model the effects of condition (atypical variation vs. consistent vs. typical variation) on accuracy (proportion looking to the target object relative to the total time looking at either the target or distractor object) during the critical analysis window (300-2000ms following target word onset). Age (scaled) was included as an additional fixed effect. We also included the maximal random effects structure (by-child and by-item intercepts and slopes). Beta regression was chosen to model the proportional dependent variable [CITE]. We used a zero-one-inflated model since 0s and 1s are possible values in our data [CITE]. This particular modeling approach requires a Bayesian framework to interpret. 

```{r cluster-permutation, message=FALSE, warning=FALSE, execute = FALSE}
library(eyetrackingR)
eyetrackingr.data <- eyetrackingr.data %>%
  mutate(trackloss = NA, 
         aoi = ifelse(target == 1, "target", "distracter"),
         target = ifelse(target == 1, TRUE, FALSE),
         distracter = ifelse(distracter, TRUE, FALSE)) %>% 
  distinct() 

# Step 1: Prepare the eyetrackingR data object
et_data <- eyetrackingr.data %>%
  group_by(pid, trial, time) %>%
  filter(n() == 1) %>%
  make_eyetrackingr_data(., 
                         participant_column = "pid",
                         trial_column = "trial",
                         time_column = "time",
                         trackloss_column = "trackloss",
                         aoi_columns = c("target", "distracter"),
                         treat_non_aoi_looks_as_missing = TRUE)

max.time <- data %>%
  summarize(max = mean(second_label) - mean(first_label) + 2000) %>%
  pull(max)

response_window <- subset_by_window(et_data,
                                     window_start_time = min(first.window),
                                     window_end_time = max(second.window), 
                                     rezero = FALSE)

response_window2 <- subset_by_window(et_data,
                                     window_start_time = min(second.window),
                                     window_end_time = max(second.window), 
                                     rezero = FALSE)

response_window1 <- subset_by_window(et_data,
                                     window_start_time = 300,
                                     window_end_time = 2000, 
                                     rezero = FALSE)

response_window_agg_by_sub <- make_time_window_data(response_window1, 
                                                    aois='target',
                                              predictor_columns=c('condition'),
                                                    summarize_by = "pid") %>%
  mutate(condition = factor(condition, levels = c("unnaturalv", "naturalv", "stable")), 
         id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id")


plot(response_window_agg_by_sub, predictor_columns="condition", dv = "Prop")

library(lme4)
library(lmerTest)  # for p-values

m <- lmer(Prop ~ condition + age_c + (1 | pid), data = response_window_agg_by_sub)
summary(m)

ggplot(response_window_agg_by_sub, aes(x=condition, y = Prop, color = condition)) +
  stat_summary(fun.data = mean_se, geom="pointrange")+
  coord_cartesian(ylim = c(0, 1))+
  labs(y = "Prop. of Target Looks")+
  geom_hline(yintercept = .5, color = "black")+
  scale_color_manual(values = condition.colors) + 
  theme_bw() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

response_time <- make_time_sequence_data(response_window1,
                                         time_bin_size = 10, 
                                         predictor_columns = c("condition"),
                                         aois = "target",
                                         summarize_by = "pid")

response_time <- response_time %>%
  mutate(condition = factor(condition, levels = c("unnaturalv", "stable", "naturalv")))

ggplot(response_time, aes(x = Time, y = Prop))+
  geom_ribbon(aes(xmin = 300, xmax = 2000),
  color = NA, fill = "#9982E1", alpha = 0.1) +
  geom_ribbon(aes(xmin = min(second.window), xmax = max(second.window)),
  color = NA, fill = "#9982E1", alpha = 0.1) +
  stat_smooth(method="loess", span=0.1, se=TRUE, alpha=0.3,
  color = "black", fill = "black") +
  labs(x = "Time from first target label onset (ms)", 
       y = "Accuracy\n(Proportion looking to target object)")+
  geom_hline(yintercept = 0.5, color = "black", linetype="dashed")+
  geom_vline(xintercept = 0, color = "#9982E1", size = 2) +
  geom_vline(xintercept = max.time-2000, color = "#9982E1", size = 2) + 
  scale_x_continuous(limits = c(-2500, 5000), breaks = c(-2500, 0, 2500, 5000)) + 
  coord_cartesian(ylim = c(0,1))+
  theme_classic(base_size = 20) + 
  theme(legend.position = "bottom",
        legend.title = element_blank())

response_time %>%
  ggplot(aes(x = Time, y = Prop, color = condition)) +
  geom_ribbon(aes(xmin = 300, xmax = 2000), color = NA, fill = "gray", alpha = 0.25) + 
  geom_ribbon(aes(xmin = min(second.window), xmax = max(second.window)), 
             color = NA, fill = "gray", alpha = 0.25) + 
  stat_smooth(method="loess", span=0.1, se=TRUE, aes(fill=condition), alpha=0.3) +
  labs(x = "Time from first target label onset (ms)", 
       y = "Accuracy\n(Proportion looking to target object)")+
  geom_hline(yintercept = .5, color = "black", linetype="dashed")+
  geom_vline(xintercept = 0, color = "black")+
  geom_vline(xintercept = max.time-2000, color = "black") + 
  scale_color_manual(values = condition.colors) + 
  scale_fill_manual(values = condition.colors) + 
  coord_cartesian(ylim = c(0,1))+
  theme_classic(base_size = 20) + 
  theme(legend.position = "bottom",
        legend.title = element_blank())

df_timeclust <- make_time_cluster_data(filter(response_time, condition %in% c("unnaturalv", "stable")), 
                                       test= "t.test", paired = TRUE,
                                       predictor_column = "condition", 
                                       threshold = 2)

plot(df_timeclust) +  ylab("T-Statistic") + theme_light() + 
  ggtitle("Both windows: Stable vs. Unnatural variation")
```

```{r cluster-based-permutation, execute = FALSE}
rerun.cbp = FALSE

if (rerun.cbp) {
library(foreach)
library(doSNOW)
library(parallel)
library(data.table)

# ==== CONFIG ====
n_perm <- 1000  # number of permutations
f_threshold <- qf(0.80, df1 = 2, df2 = 92)
time_var <- "time_bin"

# ==== Define helper functions ====

# Repeated-measures ANOVA F-value (using base R aov)
get_f_stat <- function(df) {
  tryCatch({
    out <- summary(aov(accuracy ~ condition + Error(pid / condition), data = df))
    out[[2]][[1]]["F value"][1]
  }, error = function(e) NA_real_)
}

# Identify largest cluster (sum of F-values over consecutive bins above threshold)
get_max_cluster_anova <- function(df, threshold = f_threshold) {
  df <- df %>%
    mutate(sig = ifelse(F_val >= threshold, "sig", "ns"),
           cluster = data.table::rleid(sig)) %>%
    filter(sig == "sig") %>%
    group_by(cluster) %>%
    summarise(start = min(!!sym(time_var)),
              stop = max(!!sym(time_var)) + 50,
              auc = sum(F_val), .groups = "drop")
  
  if (nrow(df) == 0) return(tibble(cluster = 0, start = 0, stop = 0, auc = 0))
  
  df[which.max(df$auc), ]
}

# Shuffle condition labels within each participant
permute_within_subjects <- function(data, seed) {
  set.seed(seed)
  data %>%
    group_by(pid, !!sym(time_var)) %>%
    mutate(condition = sample(condition)) %>%
    ungroup()
}

# ==== Set up parallel backend ====
n_cores <- detectCores() - 1
cl <- makeCluster(n_cores)
registerDoSNOW(cl)

pb <- txtProgressBar(max = n_perm, style = 3)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

# ==== Prepare data ====
dc = eyetrackingr.data %>%
  filter(time >= 300 & time <= max(second.window)) %>%
  mutate(accuracy = ifelse(fix == "target", 1, 0))

dc_binned <- dc %>%
  mutate(time_bin = floor(time / 50) * 50) %>%
  group_by(pid, condition, time_bin) %>%
  summarise(accuracy = mean(accuracy, na.rm = TRUE), .groups = "drop") %>%
  group_by(pid, time_bin) %>%
  filter(length(unique(condition)) == 3) %>%
  ungroup()

# ==== Run observed F-stats and max cluster ====
f_obs <- dc_binned %>%
  group_by(!!sym(time_var)) %>%
  group_modify(~ tibble(F_val = get_f_stat(.x))) %>%
  ungroup()

max_cluster_obs <- get_max_cluster_anova(f_obs)

# ==== Run permutations in parallel ====
perm_cluster_aucs <- foreach(i = 1:n_perm, .combine = c,
                             .options.snow = opts,
                             .packages = c("tidyverse", "data.table")) %dopar% {
  d_perm <- permute_within_subjects(dc_binned, seed = i)

  f_perm <- d_perm %>%
    group_by(!!sym(time_var)) %>%
    group_modify(~ tibble(F_val = get_f_stat(.x))) %>%
    ungroup()

  get_max_cluster_anova(f_perm)$auc
}

stopCluster(cl)

# ==== Compute p-value and plot ====
p_obs <- mean(perm_cluster_aucs > max_cluster_obs$auc)

# ggplot() +
#   geom_density(aes(x = perm_cluster_aucs), fill = "gray", alpha = 0.5) +
#   geom_vline(xintercept = max_cluster_obs$auc, color = "blue", linewidth = 1.2) +
#   labs(title = "Permutation-based Null Distribution (3-level ANOVA)",
#        subtitle = paste("Observed p =", round(p_obs, 3)),
#        x = "Max cluster AUC (F values)")
} else {
  max_cluster_obs <- tibble(
    cluster = 3, 
    start = 350,
    stop = 400, 
    auc = 2.92
  )
  p_obs <- 0.997
}
```
<!---A cluster-based permutation test identified a marginal cluster from `max_cluster_obs$start` to `max_cluster_obs$stop`ms (AUC = `r max_cluster_obs$auc`), but this cluster did not reach significance compared to the permutation-based null distribution (p = `p_obs`), suggesting at most a weak and transient effect of condition.--->

## Moment-by-moment attention measured by pupil size synchrony
### Comparing pupil size synchrony across conditions
Following @nencheva2021moment, we used dynamic time-warping distance [DTW; @tormene2009matching] to quantify the temporal alignment of participants' pupil responses when hearing individual exposure sentences or words (normalized for duration). This DTW measure quantifies the dissimilarity between two time series of pupil fluctuations. To transform these values into synchrony scores, we subtracted DTW distance scores from the maximum distance between two participants on each trial. To compare pupil size synchrony across conditions, we computed pairwise synchrony scores for all possible pairs of participants (for each exposure sentence or word per condition). 

```{r pupillometry-sentence-level}
sync_data <- read_csv("data/synchrony/combined/pupil-synchrony-exp23_sentence.csv") %>%
  group_by(participant, condition, word, exposure) %>%
  summarize(mean_synchrony = mean(synchrony, na.rm = TRUE), .groups = "drop") %>%
  mutate(pid = gsub("ex", "", participant)) %>%
  group_by(pid, word, condition) %>%
  summarize(mean_synchrony = mean(mean_synchrony)) %>%
  ungroup() %>%
  mutate(sync_z = datawizard::standardize(mean_synchrony))

acc_data <- read_csv("data/test-props.csv") %>%
  transmute(pid = pid,
            word = base_word, 
            condition = condition, 
            prop = prop) %>%
  group_by(pid, word, condition) %>%
  summarize(mean_accuracy = mean(prop)) %>%
  ungroup() %>%
  mutate(acc_z = datawizard::standardize(mean_accuracy))

combined_data <- left_join(acc_data, sync_data,
                           by = c("pid", "word", "condition")) %>%
  ungroup() %>%
  filter(!is.na(sync_z), !is.na(acc_z)) %>%
  mutate(condition = factor(condition, 
                            levels = c("unnaturalv", "stable", "naturalv"))) 

sync_model <- lmer(sync_z ~ condition + (1|pid), data = combined_data)
# confirmed no outliers!
# outliers <- performance::check_outliers(sync_model)
anova_result <- anova(sync_model)
eta <- effectsize::eta_squared(sync_model, partial = TRUE)

# pairwise ccomparions
emm_sync <- emmeans(sync_model, pairwise ~ condition)
pairs_sync <- emm_sync$contrasts %>% as.data.frame()
pairwise_es <- eff_size(emm_sync, sigma = sigma(sync_model), edf = df.residual(sync_model)) %>% as.data.frame()

# no outliers!
#outliers <- performance::check_outliers(sync_model)
#outlier_ids <- which(outliers)
#combined_data_clean <- combined_data[-outlier_ids, ]
```

A one-way ANOVA revealed a significant main effect of condition on pupil size synchrony [F(2, `r anova_result$DenDF`) = `r round(anova_result$"F value"[1], 2)`, _p_ = `r format.pval(anova_result$"Pr(>F)"[1], digits = 3)`].

Follow-up pairwise comparisons using Tukey's HSD correction indicated that pupil size synchrony was significantly higher in the typical variation condition compared to the atypical variation condition (ΔM = `r round(pairs_sync$estimate[2], 2)`, _p_ = `r round(pairs_sync$p.value[2], 3)`, Cohen's d = `r round(pairwise_es$effect.size[2], 2)`). No other pairwise comparisons were statistically significant (_p_s > 0.05).

```{r pupillometry-fig}
emmeans(sync_model, ~ condition) %>%
  as.data.frame() %>%
  ggplot(aes(x = condition, fill = condition, color = condition)) +
  geom_jitter(data = combined_data, 
              mapping = aes(y = sync_z),
              width = 0.2, alpha = 0.2, size = 1) +
  geom_errorbar(mapping = aes(y = emmean, ymin = lower.CL, ymax = upper.CL), width = 0.1, linewidth = 1.25) + 
  scale_color_manual(values = condition.colors) + 
  scale_fill_manual(values = condition.colors) + 
  scale_x_discrete(labels = c("Atypical variation", "Consistent", "Typical variation")) + 
  labs(
    title = "Pupil Synchrony Across Conditions",
    x = "Condition",
    y = "Mean Pupil Synchrony"
  ) +
  theme_minimal() + 
  theme(legend.position = "none")

ggplot(combined_data, aes(x = sync_z, y = acc_z, color = condition, fill = condition)) +
  geom_point(shape = 21, alpha = 0.1) + 
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~ condition) +
  scale_color_manual(values = condition.colors) + 
  scale_fill_manual(values = condition.colors) + 
  labs(
    x = "Mean Pupil Synchrony",
    y = "Mean Accuracy (Proportion Target Looks)",
    title = "Relationship Between Pupil Synchrony and Accuracy",
    subtitle = "Faceted by Condition"
  ) +
  theme_test() + 
  theme(legend.position = "none")

m_null <- lmer(acc_z ~ condition +  (1 | pid), data = combined_data)
m_sync <- lmer(acc_z ~ condition + sync_z + (1 | pid), data = combined_data)
anova(m_null, m_sync)

model.data <- test.props %>%
  mutate(id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id") %>%
  mutate(condition = relevel(as.factor(condition), ref = "unnaturalv"), 
         pid = paste0("LVar", id), 
         word = base_word) %>%
  left_join(sync_data, by = c("pid", "condition", "word"))

if (rerun) {
  m.sync <- brm(
    bf(prop ~ condition + age_c + sync_z + (1 + condition | id) + (1 + condition + age_c | base_word),
     phi ~ age_c,
     zoi ~ age_c,
     coi ~ condition + age_c),
  data = model.data,
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  iter = 4000,
  control = list(adapt_delta = 0.999)
  )
  saveRDS(m.sync, file = "models/m-sync.rds")
  
  # refit m.age with shared sync data
  common_data <- m.sync$data
  m.age.refit <- update(m.age, newdata = common_data, recompile = FALSE)
  saveRDS(m.sync, file = "models/m-age-refit.rds")
} else {
  m.sync <- readRDS("models/m-sync.rds")
  m.age.refit <- readRDS("models/m-age-refit.rds")
}
```

```{r pupillometry-word-level}
sync_data <- read_csv("data/synchrony/combined/pupil-synchrony-exp23.csv") %>%
  group_by(participant, condition, word, exposure) %>%
  summarize(mean_synchrony = mean(synchrony, na.rm = TRUE), .groups = "drop") %>%
  mutate(pid = gsub("ex", "", participant)) %>%
  group_by(pid, word, condition) %>%
  summarize(mean_synchrony = mean(mean_synchrony)) %>%
  ungroup() %>%
  mutate(sync_z = datawizard::standardize(mean_synchrony))

combined_data <- left_join(acc_data, sync_data,
                           by = c("pid", "word", "condition")) %>%
  ungroup() %>%
  filter(!is.na(sync_z), !is.na(acc_z)) %>%
  mutate(condition = factor(condition, 
                            levels = c("unnaturalv", "stable", "naturalv"))) 

sync_model <- lm(sync_z ~ condition, data = combined_data)

#outliers <- performance::check_outliers(sync_model)
#outlier_ids <- which(outliers)
#combined_data_clean <- combined_data[-outlier_ids, ]

#sync_model <- lmer(sync_z ~ condition + (1 | pid), data = combined_data)
#summary(sync_model)

anova_result <- anova(sync_model)
eta <- effectsize::eta_squared(sync_model, partial = TRUE)

# pairwise comparisons
emm_sync <- emmeans(sync_model, pairwise ~ condition)
pairs_sync <- emm_sync$contrasts %>% as.data.frame()
pairwise_es <- eff_size(emm_sync, sigma = sigma(sync_model), edf = df.residual(sync_model)) %>% as.data.frame()

model.data <- test.props.bywindow %>%
  filter(window == "first") %>%
  mutate(id = str_extract(pid, "[0-9][0-9][0-9]")) %>%
  left_join(ages, by = "id") %>%
  mutate(pid = paste0("LVar", id), 
       word = base_word) %>%
  left_join(sync_data, by = c("pid", "condition", "word")) %>%
  mutate(condition = factor(condition,
                          levels = c("unnaturalv",
                                     "stable", 
                                     "naturalv"))) %>%
  filter(!is.na(sync_z))

if (rerun) {
  m.sync <- brm(
    bf(prop ~ condition * sync_z + age_c + (1 + condition | id) + (1 + condition + age_c | base_word),
     phi ~ condition + age_c,
     zoi ~ condition + age_c,
     coi ~ condition + age_c),
  data = model.data,
  family = zero_one_inflated_beta(),
  chains = 4,
  cores = 4,
  save_pars = save_pars(all = TRUE),
  iter = 2000,
  warmup = 1000,
  seed = 20250803,
  control = list(adapt_delta = 0.999)
  )
  saveRDS(m.sync, file = "models/m-sync-word.rds")
  
  # refit m.age with shared sync data
  common_data <- m.sync$data
  m.age.refit <- update(m.age, newdata = common_data, recompile = FALSE)
  saveRDS(m.sync, file = "models/m-age-refit-word.rds")

  loo1 <- loo(m.age.refit, moment_match = TRUE, reloo = TRUE)
  loo2 <- loo(m.sync)
  comp <- loo_compare(loo1, loo2) %>%
    as_tibble() %>%
    rownames_to_column(var = "model")
  write_csv(comp, "models/compare-sync-word.csv")
} else {
  m.sync <- readRDS("models/m-sync-word.rds")
  m.age.refit <- readRDS("models/m-age-refit-word.rds")
  comp <- read_csv("models/compare-sync-word.csv")
}
 
ggplot(combined_data, aes(x = condition, y = sync_z, fill = condition, color = condition)) +
  geom_jitter(width = 0.2, alpha = 0.4, size = 1.5) +
  geom_boxplot(outlier.shape = NA, color = "black", fill = NA) +
  scale_color_manual(values = condition.colors) + 
  scale_fill_manual(values = condition.colors) + 
  scale_x_discrete(labels = c("Atypical variation", "Consistent", "Typical variation")) + 
  labs(
    title = "Pupil Synchrony Across Conditions",
    x = "Condition",
    y = "Mean Pupil Synchrony"
  ) +
  theme_minimal() + 
  theme(legend.position = "none")

newdata <- combined_data %>%
distinct(condition) %>%
  expand_grid(sync_z = seq(min(combined_data$sync_z, na.rm = TRUE),
                           max(combined_data$sync_z, na.rm = TRUE),
                           length.out = 100)) %>%
  mutate(age_c = 0)

epreds <- posterior_epred(m.sync, newdata = newdata, re_formula = NA)

# 3. Convert to long format and merge with newdata
epred_df <- as_tibble(epreds) %>%
  pivot_longer(cols = everything(), names_to = "row", values_to = "estimate") %>%
  mutate(row = as.integer(gsub("V", "", row))) %>%
  left_join(newdata %>% mutate(row = row_number()), by = "row")

pred_summary <- epred_df %>%
  group_by(condition, sync_z) %>%
  summarise(
    mean = mean(estimate),
    lower = quantile(estimate, 0.05),
    upper = quantile(estimate, 0.95),
    .groups = "drop"
  )

ggplot(combined_data, aes(x = sync_z, y = mean_accuracy, color = condition, fill = condition)) +
  geom_point(shape = 21, alpha = 0.1) +
  geom_line(data = pred_summary, aes(y = mean), size = 1) +
  geom_ribbon(data = pred_summary, aes(y = mean, ymin = lower, ymax = upper), alpha = 0.2, color = NA) +
  facet_wrap(~ condition) +
  scale_color_manual(values = condition.colors) +
  scale_fill_manual(values = condition.colors) +
  labs(
    x = "Mean Pupil Synchrony",
    y = "Mean Accuracy (Proportion Target Looks)",
    title = "Model-Based Relationship Between Pupil Synchrony and Accuracy",
    subtitle = "Predicted from zero-one inflated beta regression, faceted by condition"
  ) +
  theme_test() +
  theme(legend.position = "none")
```

We also examined pupil size synchrony at the word level. A one-way ANOVA revealed a significant main effect of condition on pupil size synchrony \[*F*(2, 314) = `r format(round(anova_result$"F value"[1], 2), nsmall = 2)`, *p* = `r format(round(anova_result$"Pr(>F)"[1], 3), nsmall = 3)`\]. Follow-up pairwise comparisons using Tukey's HSD correction indicated that, relative to the consistent condition, pupil size synchrony was significantly lower in the atypical variation condition condition (ΔM = `r round(pairs_sync$estimate[1], 2)`, *p* = `r round(pairs_sync$p.value[1], 3)`, Cohen's *d* = `r format(round(pairwise_es$effect.size[1], 2), nsmall = 2)`) and marginally lower in the typical variation condition (ΔM = `r round(-pairs_sync$estimate[3], 2)`, *p* = `r round(pairs_sync$p.value[3], 3)`, Cohen's *d* = `r round(pairwise_es$effect.size[3], 2)`). There was no difference in synchrony between the atypical variation condition and the typical variation condition (*p* \> 0.05).

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# Funding
This research was supported by a XXX awarded to CLW (XXX) and a National Science Foundation Graduate Research Fellowship awarded to KC.

# Acknowledgements
We are grateful to the participating families as well as the members of the Princeton Baby Lab for their feedback on this work.
