---
title: "Facilitative effects of label variation on young children's word learning"
shorttitle: "Effects of label variation on word learning"
author:
  - name: Kennedy Casey
    corresponding: true
    orcid: 0000-0001-5326-3080
    email: kcasey@princeton.edu
    affiliations:
      - name: Princeton University
        department: Department of Psychology
        address: 40 Woodlands Way
        city: Princeton
        region: NJ
        country: USA
        postal-code: 08544
  - name: Casey Lew-Williams
    orcid: 0000-0002-8781-4458
    email: caseylw@princeton.edu
    affiliations:
      - name: Princeton University
        department: Department of Psychology
        address: 40 Woodlands Way
        city: Princeton
        region: NJ
        country: USA
        postal-code: 08544
author-note:
  status-changes: 
    affiliation-change: ~
    deceased: ~
  disclosures:
    study-registration: ~
    data-sharing: ~
    related-report: ~
    conflict-of-interest: ~
    financial-support: ~
    gratitude: ~
    authorship-agreements: ~
abstract: "This document is a template."
keywords: [word learning, child-directed speech, wordplay, eye-tracking]
bibliography: LexVar.bib
format:
  apaquarto-docx: default
  apaquarto-html: default
  apaquarto-pdf:
    documentmode: man
execute: 
  echo: false
  warning: false
  message: false
  error: false
  eval: true

---
```{r setup}
library(tidyverse)
library(readxl)
library(english)
library(tools)
library(flextable)
```

# Method
Pre-registered ([(https://osf.io/uqz6d)](https://osf.io/uqz6d))

```{r, echo=FALSE}
log <- read_xlsx("data/metadata/ANON-ptcp-log.xlsx")
final.sample <- filter(log, usable_checked == "y") 
usable <- final.sample %>% nrow()
unusable <- filter(log, usable_checked == "n") %>% nrow()
unusable.lang <- filter(log, str_detect(exclusion_reason, "did not")) %>% nrow()
unusable.cal <- filter(log, str_detect(exclusion_reason, "calibration")) %>% nrow()
unusable.data <- filter(log, str_detect(exclusion_reason, "trial")) %>% nrow()
unusable.inter <- filter(log, str_detect(exclusion_reason, "interruption")) %>% nrow()

ethnicity <- final.sample %>%
  mutate(ethnicity = case_when(
    ethnicity %in% c("black", "asian", "not reported") ~ str_to_sentence(ethnicity), 
    ethnicity == "white" ~ "Non-Latinx white", 
    ethnicity == "white/hispanic" ~ "Latinx white", 
    ethnicity %in% c("white/black", "asian/white", "mixed") ~ "Mixed race"
  )) %>%
  group_by(ethnicity) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(prop = round(n/sum(n)*100, 2)) %>%
  arrange(desc(prop))
```
## Participants
Participants (_N_ = `r usable`, age range = 18--36 months, *M*~age~ = `r format(round(mean(final.sample$age), 2), nsmall = 2)`) were full-term, monolingual English-hearing children (>80% reported exposure) with no known hearing, vision, language, and/or developmental delays. Caregivers self-reported race/ethnicity information for their children in an open-ended format when they initially signed up for our volunteer database. `r sub("^([a-z])", "\\U\\1", as.character(english(ethnicity[2,]$n)), perl = TRUE)` (`r ethnicity[2,]$prop`% of the final sample) caregivers chose not to report this information. Participants were primarily `r ethnicity[1,]$ethnicity` (_N_ = `r ethnicity[1,]$n`, `r ethnicity[1,]$prop`%) with some representation of other races/ethnicities: Mixed race (_N_ = `r ethnicity[2,]$n`, `r ethnicity[2,]$prop`%), Asian (_N_ = `r ethnicity[4,]$n`, `r ethnicity[4,]$prop`%), Black (_N_ = `r ethnicity[5,]$n`, `r ethnicity[5,]$prop`%), `r ethnicity[6,]$n` (_N_ = `r ethnicity[6,]$n`, `r ethnicity[6,]$prop`%).

The final sample size was pre-registered and determined using an a priori power analysis conducted in G*Power [@faul2009statistical] and informed by prior eye-tracking studies with a similar age range [@nencheva2021moment; @quam2023protracted]. `r sub("^([a-z])", "\\U\\1", as.character(english(unusable)), perl = TRUE)` additional children participated but were excluded from the final analyses due to failure to meet the language requirement (_N_ = `r unusable.lang`), refusal to wear sticker or complete calibration (_N_ = `r unusable.cal`), significant interruption by family member (_N_ = `r unusable.inter`), or contribution of too few usable trials based on pre-registered criteria (_N_ = `r unusable.data`).

## Stimuli
A female native English speaker recorded the auditory stimuli using child-directed prosody in a sound-attenuated room.

```{r stimuli-table}
tribble(
  ~Standard, ~Redup_Typical, ~Redup_Atypical, ~Dim_Typical, ~Dim_Atypical,
  "tanzer\n/tæn.zɚ/", "tan-tan\n/tæn.tæn/", "zer-zer\n/zɚ.zɚ/", "tanny\n/tæn.i/", "y-zer\n/i.zɚ/",
  "bogdum\n/bɑg.dʌm/", "bog-bog\n/bɑg.bɑg/", "dum-dum\n/dʌm.dʌm/", "boggy\n/bɑg.i/", "y-dum\n/i.dʌm/",
  "kaisol\n/kaɪ.sɔl/", "kai-kai\n/kaɪ.kaɪ/", "sol-sol\n/sɔl.sɔl/", "kaisy\n/kaɪ.si/", "y-sol\n/i.sɔl/",
  "mobet\n/moʊ.bɪt/", "mo-mo\n/moʊ.moʊ/", "bet-bet\n/bɪt.bɪt/", "moby\n/moʊ.bi/", "y-bet\n/i.bɪt/",
  "woifer\n/wɔɪ.fɚ/", "woi-woi\n/wɔɪ.wɔɪ/", "fer-fer\n/fɚ.fɚ/", "woify\n/wɔɪ.fi/", "y-fer\n/i.fɚ/",
  "voopin\n/vu.pɪn/", "voo-voo\n/vu.vu/", "pin-pin\n/pɪn.pɪn/", "voopy\n/vu.pi/", "y-pin\n/i.pɪn/"
) %>% flextable() |>
  set_table_properties(width = 1, layout = "autofit") |>
  fontsize(size = 10, part = "all") |>
  align(align = "left", part = "all")
```

The visual stimuli consisted of six isoluminant objects with 12 unique colors (two per object) that were modeled after @nencheva2021moment's stimuli and hand-drawn using Procreate software [@procreate].

## Procedure
Toddlers sat on their caregiver's lap, approximately 60cm away from an EyeLink 1000 Plus eye-tracker, in a dimly-lit experiment room. Caregivers wore opague sunglasses to prevent them from influencing their children's behavior. We used a simple 3-point calibration procedure.

The experiment consisted of 3 blocks (one for each condition). 

During the training trials (4 per block), children saw

During the test trials (4 per block)

## Data analysis

# Results

# Discussion

# Conclusion

# References

<!-- References will auto-populate in the refs div below -->

::: {#refs}
:::

# Funding
This research was supported by a XXX awarded to CLW (XXX) and a National Science Foundation Graduate Research Fellowship awarded to KC. Additional personnel support was provided by the Princeton University Office of Undergraduate Research.

# Acknowledgements
We are grateful to the participating families as well as the members of the Princeton Baby Lab for their feedback on this work.
